# Spatial features extraction

## Hausdorff Distance (dH(X,Y))
We computed drawing error based on a shape-matching approach known as the Hausdorff distance.First, let $X$ and $Y$ be two-non empty subsets of a metric space $(M, d)$ where $M$ is set and $d$ is a metric or distance on $M$. The formula of the Hausdorff distance $d_H(X,Y)$ is as follow:
$$d_H(X,Y) = \max\{\sup_{x\in X} \inf_{y\in Y} d(x,y), \sup_{y\in Y}\inf_{x\in X} d(x,y)\}$$
From Creagh et al., 2020, it was stated that ``This metric compares the maximum distance of one set to the nearest point in another set, which can be used as a basis to compute the error between the reference way-points (interpolated into a reference shape scaled to the number of pixels drawn) and the subject’s drawing attempt". Thus we took similar approach by first interpolating the reference points to the size of the drawing attempt prior to calculating the $d_H(X,Y)$.

```{r  message=FALSE, warning=FALSE, include=FALSE}

## Function to set up the data frame to later calculate the Hausdorff distances
HausD_dataframe <- function(data, xs, ys){
  ## data is the dataframe containing drawn x and y coordinates
  ## xs and ys are the x and y coordinates
  
  appendage <- data$appendage[1]
  dl <- data$difficulty_level[1]
  spiral_dl <- GenerateSpiral(appendage,dl)
  
  ## interpolate the reference spiral to the size of x coordinate
  # Drawn coordinates
  x <- data[, xs]
  y <- data[, ys]
  dat <- fdata(data.frame(x1=x,y1=y)) ## convert to functional data class
  # reference coordinate interpolated to the lengh of x
  ref_x <- spline(spiral_dl$x, n=length(x))$y
  ref_y <- spline(spiral_dl$y, n=length(y))$y
  ref_dat <- fdata(data.frame(x2=ref_x,y2=ref_y)) ## convert to functional data class
  
  
  return(list(drawing_df = dat, ref_df = ref_dat))
}

####

HausData_func <- function(data, xs, ys){
  ## data is the entire dataset with x and y scaled
  ## xs is the x coordinates to use (either x_scf or x_scc)
  ## ys is the y coordinates to use (either y_scf or y_scc)
  
  #### intialize an empty 0 column
  data[,"hausD"] <- 0 ##  Maximum H_d
  data[,"hausDError"] <- 0 ## Sum of  H_d (error)
  data[, "hausDt"] <- 0 ## H_d normalized by time taking to complete the drawing
  data[,"hausDIqr"] <- 0 ## Interquartile range of H_d
  data[, "hausD25"] <- 0 ## Beginning error (25%)
  data[, "hausD75"] <- 0 ## Ending error (75%)
  data[,"hausDmiddle"] <- 0 ## H_d from the middle (15% to 85%)
  data[,"hausDmiddlet"] <- 0 ## Middle H_d(15% to 85%) weighted by time to completion

  
  
  # create an empty dataframe to be used later
  new.tr.level <- data[0,]
  
  new.tr.level <- new.tr.level %>% 
    # dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i))
    dplyr::select(-c(t,time,x,y,d_t,v_i,rv_i,av_i)) ## remove y and x scale
  
  ## factor levels of patient ID
  p.id <- levels(factor(data$patient_id))
  
  for (id in p.id) {
    patient_data <- subset(data, patient_id==id) ## select a single patient
    n.test <- levels(factor(patient_data$ntest_date))
    for (tn in n.test) {
      test_date <- subset(patient_data, ntest_date==tn) ## select one test date
      d.levels <- levels(factor(test_date$difficulty_level))
      for (dl in d.levels) {
        d.level <- subset(test_date, difficulty_level==dl) ## select difficulty level
        tr.levels <- levels(factor(d.level$trial_id))
        #ct <- 0 ## Intialize count to be used in the trial level
        for (tr in tr.levels){
          tr.level <- subset(d.level, trial_id==tr) ## select trial_id
          dim_tr <- dim(tr.level)[1] ## Make sure the dimension of each trial > 2
          if (dim_tr>5){
            print(c(id,tn,dl,tr)) ## to print where we are in the loop
            #ct = ct + 1 ## start a count for increment
           #print(ct) ## to check the loop is working fine 
            
            ## Get the drawing and reference data
            dat <- HausD_dataframe(tr.level, xs, ys)$drawing_df$data
            ref_dat <- HausD_dataframe(tr.level, xs, ys)$ref_df$data
            ## Get the total time use to complete th drawing
            time_n <- tr.level$time[dim_tr]
            
            ## Subset data from begining 25%, ending 75%, and middle 15%-85%
            dat25 <- fdata(dat[1:round(dim_tr*0.25),])
            dat75 <- fdata(dat[round(dim_tr*0.75):dim_tr,])
            dat15_85 <- fdata(dat[round(dim_tr*0.15):round(dim_tr*0.85),])
            
            dim25 <- dim(dat[1:round(dim_tr*0.25),])[1] ## Number of touchpoints at the beginning
            dim75 <- dim(dat[round(dim_tr*0.75):dim_tr,])[1] ## Number of touchpoints at the end
              
            ref_dat25 <- fdata(ref_dat[1:round(dim_tr*0.25),])
            ref_dat75 <- fdata(ref_dat[round(dim_tr*0.75):dim_tr,])
            ref_dat15_85 <- fdata(ref_dat[round(dim_tr*0.15):round(dim_tr*0.85),])
            
            
            #### Calculate the different metrics
            tr.level[tr.level$trial_id==tr,"hausD"] <- 
              max(metric.hausdorff(fdata(dat),fdata(ref_dat)), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDError"] <- 
              sum(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDt"] <-
              max(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)/time_n
            tr.level[tr.level$trial_id==tr,"hausDIqr"] <-
              IQR(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausD25"] <- 
              max(metric.hausdorff(dat25,ref_dat25), na.rm = TRUE)/dim25
            tr.level[tr.level$trial_id==tr,"hausD75"] <- 
              max(metric.hausdorff(dat75,ref_dat75), na.rm = TRUE)/dim75
            tr.level[tr.level$trial_id==tr,"hausDmiddle"] <-
              max(metric.hausdorff(dat15_85,ref_dat15_85), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDmiddlet"] <-
              max(metric.hausdorff(dat15_85,ref_dat15_85), na.rm = TRUE)/time_n
            
            ## Remove some variables and take first row only
            tr.level <- tibble::as_tibble(tr.level) %>% 
              # dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i))
              dplyr::select(-c(t,time,x,y,d_t,v_i,rv_i,av_i)) %>% ## remove s and y scale
              group_by(patient_id, ntest_date, difficulty_level, trial_id) %>% 
              slice(1)
            
            tr.level <- as.data.frame(tr.level) ## convert to data.frame
            
            new.tr.level <- rbind(new.tr.level, tr.level) ## append the data
             
          }
        }
      }
    }
  }
  
  return(as.data.frame(new.tr.level))  
}


```






```{r  message=FALSE, warning=FALSE, include=FALSE}

# #### calculate data frame with hausdorff distance metrics
# 
# ## Coordinates x and y
# ptm <- proc.time() ## Take 56.85 min
# ms_data.hausD_xy <- HausData_func(ms_data4, "x", "y")
# proc.time() - ptm
# #
# write.csv(ms_data.hausD_xy, file = "ms_data.hausD_xy.csv")
# 
# 
# ## Coordinates x_scf and y_scf
# ptm <- proc.time() ## Take 59 min
# ms_data.hausD_xscf <- HausData_func(ms_data4, "x_scf", "y_scf")
# proc.time() - ptm
# 
# 
# write.csv(ms_data.hausD_xscf, file = "ms_data.hausD_xscf.csv")
# 
# 
# ## Coordinates x_scc and y_scc
# ptm <- proc.time() ## Take 59 min
# ms_data.hausD_xscc <- HausData_func(ms_data4, "x_scc", "y_scc")
# proc.time() - ptm
# #
# #
# write.csv(ms_data.hausD_xscc, file = "ms_data.hausD_xscc.csv")

ms_data.hausD_xy <- read.csv("ms_data.hausD_xy.csv", header = TRUE)
ms_data.hausD_xscf <- read.csv("ms_data.hausD_xscf.csv", header = TRUE)
ms_data.hausD_xscc <- read.csv("ms_data.hausD_xscc.csv", header = TRUE)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}
dl <- c("Difficulty Level 1", "Difficulty Level 2", "Difficulty Level 3")
ms_data.hausD_xy$difficulty_level <- factor(ms_data.hausD_xy$difficulty_level, labels = dl)
ms_data.hausD_xscf$difficulty_level <- factor(ms_data.hausD_xscf$difficulty_level, labels = dl)
ms_data.hausD_xscc$difficulty_level <- factor(ms_data.hausD_xscf$difficulty_level, labels = dl)

p_group <- "Others"
##### Plot density of the Hausdorff distances Error
## raw drawing
ggplot(data = subset(ms_data.hausD_xy, diagnosis_group == p_group), aes(x = hausDError)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  geom_vline(data = plyr::ddply(subset(ms_data.hausD_xy, diagnosis_group == p_group),
                     "difficulty_level", summarize, m.error= mean(hausDError)), 
             aes(xintercept = m.error), color="blue", linetype="dashed", size=1.2) +
  facet_wrap(~difficulty_level, scales = "free_y", nrow = 3) +
  #scale_x_continuous(labels = scales::scientific) +
  labs(x = "HausDError", y = "Count") + 
  theme_bw() + Nice.Label

## Normalized by first point rule
ggplot(data = subset(ms_data.hausD_xscf,diagnosis_group == p_group), aes(x = hausDError)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  geom_vline(data = plyr::ddply(subset(ms_data.hausD_xy, diagnosis_group == p_group),
                     "difficulty_level", summarize, m.error= mean(hausDError)), 
             aes(xintercept = m.error), color="blue", linetype="dashed", size=1.2) +
  facet_wrap(~difficulty_level, scales = "free_y", nrow = 3) +
  labs(x = "HausDError", y = "Count") + 
  theme_bw() + Nice.Label

## spiral normalized drawing using center rule
ggplot(data = subset(ms_data.hausD_xscc, diagnosis_group == p_group), aes(x = hausDError)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  geom_vline(data = plyr::ddply(subset(ms_data.hausD_xy, diagnosis_group == p_group),
                     "difficulty_level", summarize, m.error= mean(hausDError)), 
             aes(xintercept = m.error), color="blue", linetype="dashed", size=1.2) +
  facet_wrap(~difficulty_level, scales = "free_y", nrow = 3) +
  labs(x = "HausDError", y = "Count") + 
  theme_bw() + Nice.Label

```



## Drawing Error using Trapezoidal method and 2D MSE/RMSE

Here we calculated two different type of errors. The first error was calculated using the trapezoidal rule to integrate over the two spiral regions. The error was calculated by finding the intersection of the two spiral region such as the difference between the two area. We first note from [Aghanavesi et al., 2017](https://www.sciencedirect.com/science/article/pii/S2352914817300230) that the trapezoidal formula is
$$\int_{x_n}^{x_{n+1}} f(x) dx = \frac{b-a}{2N}\sum_{n=1}^{N}[f(x_n)-f(x_{n+1})]$$
where $N$ is the total number of points and $\frac{b-a}{2}$ is the spacing between the points. Suppose the reference spiral and the patients spiral are denoted by the functions $f_{ref}(x,y)$ and $f_{pat}(x,y)$ respectively. Then the error based on the trapezoidal rule becomes
$$Error = \left| \int_{x_n}^{x_{n+1}} f_{ref}(x) dx - \int_{x_n}^{x_{n+1}} f_{pat}(x) dx \right|$$
where $|.|$ is the abosulte value of the difference between the two Area Under the Curve (AUC). We now proceed with the second type of errors.


The Mean-Squared Error (MSE) between two images $f_{ref}(x,y)$ (reference image) and $f_{pat}(x,y)$ can be described as follow from [website](http://homepages.inf.ed.ac.uk/rbf/CVonline/LOCAL_COPIES/VELDHUIZEN/node18.html) and [Asamoah et al., 2018](https://www.ijcaonline.org/archives/volume181/number22/asamoah-2018-ijca-917899.pdf):

$$MSE = \frac{1}{M\times N}\sum_{n=1}^{M}\sum_{m=1}^{N}\left[f_{ref}(x,y) - f_{pat}(x,y)\right]^2$$


where $M$ and $N$ are the width and height of the images. In our case, we are in 2-D and thus $N=2$ while $M$ varies according to the trial level. We first use Spline interpolation to get the reference image to the same length as the patient drawing. 

We also calculated ``center of the shoot" as the Euclidean distance by which the touch coordinates misses the centre of the shape following the formula:

$$Center_{Sh}((x_1,y_1),(x_2,y_2)) = \sqrt{(x_1-x_2)^2 + (y_1-y_2)^2}$$


We continued by calculating the correlation coefficient using the following 2D correlation coefficient formula from [Aljanabi et al., 2018](https://www.hindawi.com/journals/mpe/2018/9801308/):

$$r = \frac{\sum_{m=1}^{M}\sum_{n=1}^{N} (A_{MN} - \bar{A})(B_{MN} - \bar{B})}{\sqrt{\left(\sum_{m=1}^{M}\sum_{n=1}^{N} (A_{MN} - \bar{A})^2\right)\left(\sum_{m=1}^{M}\sum_{n=1}^{N} (B_{MN} - \bar{B})^2\right)}}$$
where $A_{MN}$ and $B_{MN}$ are the image coordinates points with dimension $M\times 2$ given that we have $x$ and $y$ coordinate. $\bar{A} = \frac{\sum_{i}x_i +\sum_{i}y_i}{2M}$ and $\bar{B}$ following similar formula are the image mean.



```{r  message=FALSE, warning=FALSE, include=FALSE}

## Function to calculate area under the curve using trapezoidal rule

####

Error_func <- function(data, xs, ys){
  ## data is the entire dataset with x and y scaled
  ## xs is the x coordinates to use (either x_scf or x_scc)
  ## ys is the y coordinates to use (either y_scf or y_scc)
  
  #### intialize an empty 0 column
  data[,"ErrAUC"] <- 0 ##  AUC error
  data[,"MSE"] <- 0 ##  MSE
  data[,"RMSE"] <- 0 ##  RMSE
  data[, "Center_Sh"] <- 0 ## Center of Shoot
  data[,"t_final"] <- 0 ## time taken to complete the drawing
  data[,"total_asym"] <- 0 ## total asymmetry
  data[,"true_asym"] <- 0 #3 difference between patient and reference asymmetry
  data[,"Corr"] <- 0 ## 2D cross-correlation between the two images
  data[,"ImEntropy_pat"] <- 0 ## image entropy of shape drawn
  data[,"ImEntropy_ratio"] <- 0 ## image entropy of reference shape divide by reference 
  
  

  
  
  # create an empty dataframe to be used later
  new.tr.level <- data[0,]
  
  new.tr.level <- new.tr.level %>% 
   # dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i))
    dplyr::select(-c(t,time,x,y,d_t,v_i,rv_i,av_i)) # to remove x and y scale
  
  ## factor levels of patient ID
  p.id <- levels(factor(data$patient_id))
  
  for (id in p.id) {
    patient_data <- subset(data, patient_id==id) ## select a single patient
    n.test <- levels(factor(patient_data$ntest_date))
    for (tn in n.test) {
      test_date <- subset(patient_data, ntest_date==tn) ## select one test date
      d.levels <- levels(factor(test_date$difficulty_level))
      for (dl in d.levels) {
        d.level <- subset(test_date, difficulty_level==dl) ## select difficulty level
        tr.levels <- levels(factor(d.level$trial_id))
        for (tr in tr.levels){
          tr.level <- subset(d.level, trial_id==tr) ## select trial_id
          dim_tr <- dim(tr.level)[1] ## Make sure the dimension of each trial > 2
          if (dim_tr>2){
            
            ## generate the reference spiral
            appendage <- tr.level$appendage[1]
            dl <- tr.level$difficulty_level[1]
            spiral_dl <- GenerateSpiral(appendage,dl)
            
            ## Create a spline interpolation to the length of tr.level
            x_ref <- spline(spiral_dl$x, n = dim_tr)$y
            y_ref <- spline(spiral_dl$y, n = dim_tr)$y
            ## Extract x and y coordinate of the patients drawing
            x_pat <- tr.level[, xs]
            y_pat <- tr.level[, ys]
            
            ## Center of reference point
            x_ref0 <- x_ref[1]
            y_ref0 <- y_ref[1]
            ## Center of patient drawing
            if (tr.level[1, ys] > tr.level[dim_tr, ys]){
              x_pat0 <- tr.level[dim_tr, xs]
              y_pat0 <- tr.level[dim_tr, xs]
            } else if (tr.level[1, ys] < tr.level[dim_tr, ys]) {
              x_pat0 <- tr.level[1, xs]
              y_pat0 <- tr.level[1, xs]
            } else {
              x_pat0 <- 0
              y_pat0 <- 0
             }

            
            ## Calculate the Area under the curve
            AUC_ref <- AUC(x = x_ref, y = y_ref, method = "trapezoid")
            AUC_patient <- AUC(x = x_pat, y = y_pat, method = "trapezoid")
            
            ## Calculate the sum of square in x and y coordinates
            x_sq <- sum((x_ref-x_pat)^2, na.rm = TRUE)
            y_sq <- sum((y_ref-y_pat)^2, na.rm = TRUE)
            
            ## Reference and patient asymmetry
            ref_asym <- 
              abs(abs(min(x_ref)) - abs(max(x_ref)))/(abs(min(x_ref)) + abs(max(x_ref))) +
              abs(abs(min(y_ref)) - abs(max(y_ref)))/(abs(min(y_ref)) + abs(max(y_ref)))
            pat_asym <- 
              abs(abs(min(x_pat)) - abs(max(x_pat)))/(abs(min(x_pat)) + abs(max(x_pat))) +
              abs(abs(min(y_pat)) - abs(max(y_pat)))/(abs(min(y_pat)) + abs(max(y_pat)))
            
            ### Correlation metrics
            ## get the reference and patient matrix of x and y coordinates
            mat_ref <- as.matrix(data.frame(x_ref,y_ref))
            mat_pat <- as.matrix(data.frame(x_pat,y_pat))
            
            ## calculate the mean of the coordinates in both matrix
            m_ref_mean <- mean(as.vector(mat_ref), na.rm = TRUE)
            m_pat_mean <- mean(as.vector(mat_pat), na.rm = TRUE)
            
            #calculate the numerator of the correlation coefficient
            r_num <-sum((mat_ref - m_ref_mean)*(mat_pat - m_pat_mean), na.rm = TRUE)
            # calculate the denominator parts under the quareroot
            r_denom <- sum((mat_ref - m_ref_mean)^2, na.rm = TRUE)*sum(
              (mat_pat - m_pat_mean)^2, na.rm = TRUE) 
            
            #### Image entropy
            ## 2D dsicretization 
            bin_ref <- discretize2d(x_ref,y_ref,numBins1 = 50,numBins2 = 50)
            bin_pat <- discretize2d(x_pat,y_pat,numBins1 = 50,numBins2 = 50)
            
            
            
              

            
            #### Calculate the different metrics
            tr.level[tr.level$trial_id==tr,"ErrAUC"] <- abs(AUC_ref - AUC_patient)
            tr.level[tr.level$trial_id==tr,"MSE"] <- 
              sum(x_sq, y_sq, na.rm = TRUE)/(2*dim_tr) ## 2 for x and y
            tr.level[tr.level$trial_id==tr,"RMSE"] <- 
              sqrt(sum(x_sq, y_sq, na.rm = TRUE)/(2*dim_tr))
            tr.level[tr.level$trial_id==tr,"Center_Sh"] <- 
              sqrt((x_ref0 - x_pat0)^ 2 + (y_ref0 - y_pat0)^2)
            tr.level[tr.level$trial_id==tr,"t_final"] <- tr.level$time[dim_tr]
            tr.level[tr.level$trial_id==tr,"total_asym"] <- pat_asym
            tr.level[tr.level$trial_id==tr,"true_asym"] <- abs(pat_asym - ref_asym)
            tr.level[tr.level$trial_id==tr, "Corr"] <- r_num/sqrt(r_denom)
            tr.level[tr.level$trial_id==tr, "ImEntropy_pat"] <- shannonX(bin_pat)$shannon
            tr.level[tr.level$trial_id==tr, "ImEntropy_ratio"] <- 
              (shannonX(bin_pat)$shannon)/(shannonX(bin_ref)$shannon)

            
            ## Remove some variables and take first row only
            tr.level <- tibble::as_tibble(tr.level) %>% 
              # dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i))
              dplyr::select(-c(t,time,x,y,d_t,v_i,rv_i,av_i)) %>%
              group_by(patient_id, ntest_date, difficulty_level, trial_id) %>% 
              slice(1)
            
            tr.level <- as.data.frame(tr.level) ## convert to data.frame
            
            new.tr.level <- rbind(new.tr.level, tr.level) ## append the data
             
          }
        }
      }
    }
  }
  
  return(as.data.frame(new.tr.level))  
}


```




```{r  message=FALSE, warning=FALSE, include=FALSE}

#### calculate data frame with Area Under the curve

## Coordinates x and y
ptm <- proc.time() ## Take 2.933 min
ms_data.Error_xy <- Error_func(ms_data4, "x", "y")
proc.time() - ptm
#


## Coordinates x_scf and y_scf
ptm <- proc.time() ## Take 2.98 min
ms_data.Error_xscf <- Error_func(ms_data4, "x_scf", "y_scf")
proc.time() - ptm


## Coordinates x_scc and y_scc
ptm <- proc.time() ## Take 2.78 min
ms_data.Error_xscc <- Error_func(ms_data4, "x_scc", "y_scc")
proc.time() - ptm
#

```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}

## Plot AUC Error

ms_data.Error_xy$difficulty_level <- factor(ms_data.Error_xy$difficulty_level, labels = dl)
ms_data.Error_xscf$difficulty_level <- factor(ms_data.Error_xscf$difficulty_level, labels = dl)
ms_data.Error_xscc$difficulty_level <- factor(ms_data.Error_xscf$difficulty_level, labels = dl)

##### Plot density of the AUC Error
## raw drawing
ggplot(data = ms_data.Error_xy, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label

## Normalized by first point rule
ggplot(data = ms_data.Error_xscf, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label

## spiral normalized drawing using center rule
ggplot(data = ms_data.Error_xscc, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label


## Ground truth data--9HPT average
ggplot(data = ms_data.Error_xscc, aes(x = X9HPT.Avg)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "9HPT Average", y = "Count") + 
  theme_bw() + Nice.Label

```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}

#### Plot MSE Error
## raw drawing
ggplot(data = ms_data.Error_xy, aes(x = MSE)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Mean Square Error", y = "Count") + 
  theme_bw() + Nice.Label

## Normalized by first point rule
ggplot(data = ms_data.Error_xscf, aes(x = MSE)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Mean Square Error", y = "Count") + 
  theme_bw() + Nice.Label

## spiral normalized drawing using center rule
ggplot(data = ms_data.Error_xscc, aes(x = MSE)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Mean Square Error", y = "Count") + 
  theme_bw() + Nice.Label

```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}


##### Plot density of the AUC Error
## raw drawing
ggplot(data = ms_data.Error_xy, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label

## Normalized by first point rule
ggplot(data = ms_data.Error_xscf, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label

## spiral normalized drawing using center rule
ggplot(data = ms_data.Error_xscc, aes(x = ErrAUC)) +
  geom_histogram(bins = 35, fill="darkgray", color="black") +
  facet_wrap(~difficulty_level, scales = "free", nrow = 3) +
  labs(x = "Drawing Error Using AUC", y = "Count") + 
  theme_bw() + Nice.Label

```


## Approximate Entropy (ApEn)


```{r  message=FALSE, warning=FALSE, include=FALSE}

### Approximate Entropy (ApEn)
## Interpolate all v_i, r_vi, av_i over a fixed length L = 500 (see similar approach in Creagh et al., 2020) and calculate their entropy 
ms_data.apen <- ms_data4 %>% 
   dplyr::group_by(patient_id,ntest_date,difficulty_level, trial_id) %>%
  dplyr::summarise(v_apen = ApEn(spline(v_i, n = 500)$y), 
            rv_apen = ApEn(spline(rv_i, n = 500)$y), 
            av_apen = ApEn(spline(av_i, n = 500)$y))


p1.apen <- subset(ms_data.apen, patient_id=="NIB632" & ntest_date=="1")
p1.apen <- p1.apen %>% select(patient_id, difficulty_level, v_apen)
p1.apen$difficulty_level <- as.factor(p1.apen$difficulty_level)
  
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

####

p1.apen.sum <- data_summary(p1.apen, "v_apen", "difficulty_level")
  
ggplot(p1.apen.sum, aes(x=difficulty_level, y=v_apen)) + 
    geom_bar(stat="identity", color="black", colour = "gray", width = 0.5, 
           position=position_dodge()) +
  labs(x = "Difficulty Levels", y = "Approximate Speed Entropy") +
    geom_errorbar(aes(ymin=v_apen-sd, ymax=v_apen+sd), width=.2, size = 0.9,
                  position=position_dodge(0.05))+ theme_bw() + Nice.Label


###

```

