# Introduction

Smartphone devices may be an easier alternative to obtain data from patients outside of clinics and hospitals environment. In this project, smartphone-based data were obtained from Multiple Sclerosis (MS) patients using a drawing a spiral test. The test was administered in order to measure upper extremity functionality. 

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
### Libraries
library(openxlsx) ## v. 4.1.5
library(dplyr) ## v. 1.0.0
library(ggplot2) ## v. 3.3.1
library(wavelets) ## v. 0.3.0.2 ## for wavelet transform -- https://cran.r-project.org/web/packages/wavelets/wavelets.pdf
library(signal) ## v. 0.7.6 ## for filtering -- https://cran.r-project.org/web/packages/signal/signal.pdf
```




```{r  message=FALSE, warning=FALSE, include=FALSE}
### Import data
## Description of the data is as follow
# metaData.ID = Subject ID
# x: distance (in inches) of the spot drawn from the center of the screen
# y: distance (in inches) of the spot drawn from the center of the screen
# p: estimated pressure of the tap (based off of surface area)
# t: UNIX timestamp of when the drawing happened (to the millisecond)
# sumData.Num turns: the number of turns in the spiral 
# sumData.Line width: the width of the lines of the spiral
# time.point: UNIX timestamp of the drawing (this time in seconds)
# metaData.ID: Subject/patient specific ID

ms_data <- data_map <- read.xlsx("spiralLH_Obscured.xlsx", colNames = TRUE)

## add a test date to the data
ms_data$testdate <- substr(ms_data$metaData.TEST_DATE,1,10)

## add visit number based on patientID and difficulty level, remove some columns and change column names
ms_data2 <- ms_data %>% 
  group_by(metaData.ID, metaData.DIFFICULTY) %>%
  mutate(test_numb = factor(testdate, labels = 1:length(unique(testdate)))) %>% 
  select(metaData.ID,x,y,t,p,metaData.DIFFICULTY,sumData.Num.turns,
         sumData.Line.width,testdate, test_numb) %>% 
  rename(patient_id=metaData.ID,difficulty_level=metaData.DIFFICULTY,
         turn_numb = sumData.Num.turns, line_width=sumData.Line.width)
  
ms_data2$t <- ms_data2$t/1000 ## convert time point from milliseconds to seconds  
#p1 <- subset(ms_data2, patient_id=="rep067")
```


# Features extraction

As previously done in [Creagh et al., 2020](https://iopscience.iop.org/article/10.1088/1361-6579/ab8771/pdf), several features can extracted from drawing shape by an MS patient to capture temporal, spatial,and spatiotemporal factors in the drawing task that could be an indicative of manual dexterity. Thus we start by calculating each of these features:

## Temporal features

To measure temporal irregularities, several authors ( [Banaszkiewicz et al., 2008](https://europepmc.org/article/med/19353440), [Memedi et al., 2015](https://www.mdpi.com/1424-8220/15/9/23727), [Creagh et al., 2020](https://iopscience.iop.org/article/10.1088/1361-6579/ab8771/pdf) ) calculated the drawing velocities,angular and radial velocities to illustrate temporal features that may emerge in the upper extremity function of MS patients. These features were calculated using the following formula from ( [Memedi et al., 2015](https://www.mdpi.com/1424-8220/15/9/23727) and [Creagh et al., 2020](https://iopscience.iop.org/article/10.1088/1361-6579/ab8771/pdf) ):


$$v = \sum_{i=1}^{N-1} \frac{\sqrt{(x_{i+1}-x_i)^2 + (y_{i+1}-y_i)^2}}{t_{i+1}-t_i}$$
with $v$ the drawing velocity. Radial velocity (RV) is calculated as:
$$RV = \sum_{i=1}^{N-1} \frac{r_{i+1}-r_i}{t_{i+1}-t_i}$$
where $r = \sqrt{x^2+y^2}$ is the radius. Angular velocity (RHOV) is also calculated as 

$$RHOV = \sum_{i=1}^{N-1} \frac{\theta_{i+1}-\theta_i}{t_{i+1}-t_i}$$

where $\theta = tan^{-1}\left(\frac{y}{x}\right)$. $x$ and $y$ in each formula are the coordinates of the drawing pixel and $t$ is time in seconds. $N$ is the total number of pixel.




```{r  message=FALSE, warning=FALSE, include=FALSE}

### ## Initialize velocity (v), radial velocity (rv), angular velocity (rhov)
ms_data2[,"d_t"] <- 0 ## intialize an empty 0 column of delta time
ms_data2[,"v_i"] <- 0 ## intialize an empty 0 column of velocities
ms_data2[,"rv_i"] <- 0 ## intialize an empty 0 column of radial velocities
ms_data2[,"av_i"] <- 0 ## intialize an empty 0 column of angular velocities

# create an empty dataframe to be used later
new.d.level <- ms_data2[0,]

## factor levels of patient ID
p.id <- levels(factor(ms_data2$patient_id))

ptm <- proc.time() ## Check how long it will run (will take 1h 52 min to run)

for (id in p.id) {
  count = count + 1
  patient_data <- subset(ms_data2, patient_id==id) ## select a single patient
  n.test <- levels(factor(patient_data$test_numb))
  for (tn in n.test) {
    test_date <- subset(patient_data, test_numb==tn) ## select one test date
    d.levels <- levels(factor(test_date$difficulty_level))
    for (dl in d.levels) {
      d.level <- subset(test_date, difficulty_level==dl) ## select difficulty level 
      
      for (i in 1:(dim(d.level)[1] - 1)) {
        d.level[i+1,"d_t"] <- (d.level$t[i+1]-d.level$t[i]) # delta time
        d.level[i+1,"v_i"] <- sqrt((d.level$x[i+1] - d.level$x[i])^2 + 
                                     (d.level$y[i+1] - d.level$y[i])^2
                                   )/(d.level$t[i+1]-d.level$t[i]) # velocity
        d.level[i+1,"rv_i"] <- (sqrt((d.level$x[i+1])^2 + (d.level$y[i+1])^2) -                                       sqrt((d.level$x[i])^2 + (d.level$y[i])^2)
                                )/(d.level$t[i+1]-d.level$t[i]) # radial velocity
        d.level[i+1,"av_i"] <- (atan(d.level$y[i+1]/d.level$x[i+1]) - 
                                atan(d.level$y[i]/d.level$x[i])
                                )/(d.level$t[i+1]-d.level$t[i]) # angular velocity
        
      }
        new.d.level <- rbind(new.d.level, d.level) ## append the data
    }
  }
}
proc.time() - ptm

## call the final data, ms_data3
ms_data3 <- new.d.level


## Write the ms_data3 containing vi, rvi, and avi to the file to be used later
#write.xlsx(ms_data3,file = "new_spiralLH.xlsx", colNames = TRUE)

# #### Tryout for a single patient
# 
p1 <- subset(ms_data3, patient_id=="rep269") ## select single patient (rep067)
p11 <- subset(p1, test_numb=="1") ## select one test date (1)
p111 <- subset(p11, difficulty_level=="1") ## select single difficulty level (1)

### filter the speed fequency by 8 Hz
bf <- butter(3, 0.08) # 8 Hz low-pass filter
t <- p111$t - p111$t[1]
x <- p111$v_i
z <- filter(bf, x) # apply filter
plot(t, x, type = "l")
lines(t, z, col = "red")

# 
# d.t = rep(0, dim(p111)[1]) ## initialize delta time
# p111_v = rep(0, dim(p111)[1]) ## initialize the velocity of p111
# p111_rv = rep(0, dim(p111)[1]) ## initialize the radial velocity
# p111_av = rep(0, dim(p111)[1]) ## initialize the angular velocity
# 
# for (i in 1:(dim(p111)[1]-1)){
#   d.t[i+1] = (p111$t[i+1]-p111$t[i])
#   p111_v[i+1] <- sqrt((p111$x[i+1] - p111$x[i])^2 + (p111$y[i+1] - p111$y[i])^2)/(p111$t[i+1]-p111$t[i])
#   p111_rv[i+1] <-  (sqrt((p111$x[i+1])^2 + (p111$y[i+1])^2) - sqrt((p111$x[i])^2 + (p111$y[i])^2))/(p111$t[i+1]-p111$t[i])
#   p111_av[i+1] <-  (atan(p111$y[i+1]/p111$x[i+1]) - atan(p111$y[i]/p111$x[i]))/(p111$t[i+1]-p111$t[i])
# }
# 
# p111_rv.dwt <- dwt(p111_rv, filter = "d10")
# cp111_rv.dwt <- p111_rv.dwt@filter@h ## Coefficient
```










