# Aggregation of all features


```{r  message=FALSE, warning=FALSE, include=FALSE}

##### Patient derived features

##Extract feateares from Hausdorff distance datataframe
ms_data.hausD <- ms_data.hausD_xy %>% 
  dplyr::select(patient_id, ntest_date, difficulty_level, trial_id,
                hausD, hausDError, hausDt, hausDIqr, hausD25, hausD75, hausDmiddle, hausDmiddlet)
ms_data.hausD$difficulty_level <- factor(ms_data.hausD$difficulty_level, labels = c("1","2","3"))

##Extract feateares from Error datataframe
ms_data.Error <- ms_data.Error_xy %>% 
  dplyr::select(patient_id, ntest_date, difficulty_level, trial_id,
                ErrAUC, MSE, RMSE, Center_Sh, t_final, total_asym, true_asym, Corr, 
                ImEntropy_pat, ImEntropy_ratio)
ms_data.Error$difficulty_level <- factor(ms_data.Error$difficulty_level, labels = c("1","2","3"))

##Extract age, diagnosis_group, and gender from Error datataframe
ms_data.stat <- ms_data.Error_xy %>% 
  dplyr::select(patient_id, ntest_date, difficulty_level, trial_id, gender, diagnosis_group, age,
                appendage, dominant_hand)
ms_data.stat$difficulty_level <- factor(ms_data.stat$difficulty_level, labels = c("1","2","3"))
ms_data.stat$dominant_hand <- ifelse(ms_data.stat$dominant_hand=="Right","RH","LH")

## Determine if patient use dominant hand or not
ms_data.stat$dominant_hand_use <- ifelse(ms_data.stat$appendage==ms_data.stat$dominant_hand,
                                         "YES","NO")


## Merge all features
merge_cols <- c("patient_id", "ntest_date","difficulty_level","trial_id")

pat_feature <- merge(ms_data.stat, ms_data.sum, by = merge_cols, all = TRUE)
pat_feature <- merge(pat_feature, ms_data.psd, by = merge_cols, all = TRUE)
pat_feature <- merge(pat_feature, ms_data.apen, by = merge_cols, all = TRUE)
pat_feature <- merge(pat_feature, ms_data.hausD, by = merge_cols, all = TRUE)
pat_feature <- merge(pat_feature, ms_data.Error, by = merge_cols, all = TRUE)


## Change the feature name with F1 to Fn
pat_feat <- pat_feature 
feature.name <- colnames(pat_feat)[11:50]
new_feature.name <- paste(rep("F",length(feature.name)),1:length(feature.name),sep = "")

## To recognize which label is what column name
pat_label_table <- data.frame(Label = new_feature.name, Name = feature.name)

## Change the patient feature label in dataframe
colnames(pat_feature)[11:50] <- new_feature.name

```



```{r  message=FALSE, warning=FALSE, include=FALSE}

##### Clinical derived features

cl_feature <- ms_data.Error_xy %>% 
  dplyr::select(patient_id, ntest_date, difficulty_level, trial_id, 
                X9HPT.Avg, BMI, grep("CL",names(ms_data.Error_xy), value = TRUE)) 

## dominant_hand_use to cl_feature
cl_feature <- merge(ms_data.stat, cl_feature, 
                    by = c("patient_id", "ntest_date","difficulty_level", "trial_id"))

## Change the clinical feature name with C1 to Cn
cl_feat <- cl_feature 
cl.name <- colnames(cl_feat)[11:42]
new_cl.name <- paste(rep("C",length(cl.name)),1:length(cl.name),sep = "")


## To recognize which label is what column name (see chapter 1 for full name)
cl_label_table2 <- data.frame(Label = new_cl.name, Name = cl.name)

## Change the patient feature label in dataframe
colnames(cl_feature)[11:42] <- new_cl.name

```


```{r  message=FALSE, warning=FALSE, include=FALSE}
#### Extract Right and Left dominant hand from both datasets
# pat_feat.RH <- subset(pat_feature, dominant_hand=="RH")
# pat_feat.LH <- subset(pat_feature, dominant_hand=="LH")
# 
# cl_feat.RH <- subset(cl_feature, dominant_hand=="RH")
# cl_feat.LH <- subset(cl_feature, dominant_hand=="LH")
# 
# #### Extract if patient use dominant hand or not
# pat_feat.RH.Yes <- subset(pat_feat.RH, dominant_hand_use=="YES")
# pat_feat.RH.No <- subset(pat_feat.RH, dominant_hand_use=="NO")
# pat_feat.LH.Yes <- subset(pat_feat.LH, dominant_hand_use=="YES")
# pat_feat.LH.No <- subset(pat_feat.LH, dominant_hand_use=="NO")
# 
# cl_feat.RH.Yes <- subset(cl_feat.RH, dominant_hand_use=="YES")
# cl_feat.RH.No <- subset(cl_feat.RH, dominant_hand_use=="NO")
# cl_feat.LH.Yes <- subset(cl_feat.LH, dominant_hand_use=="YES")
# cl_feat.LH.No <- subset(cl_feat.LH, dominant_hand_use=="NO")


#### Extract if patient use dominant hand or not
pat_feat.dom <- subset(pat_feature, dominant_hand_use=="YES")
pat_feat.nondom <- subset(pat_feature, dominant_hand_use=="NO")


cl_feat.dom <- subset(cl_feature, dominant_hand_use=="YES")
cl_feat.nondom <- subset(cl_feature, dominant_hand_use=="NO")

```

## Identification of relevent difficulty level

In what follow, we will identify which difficulty level is most relevent for our analysis. This will be done by computing the unpaired two samples Wilcoxon test between the Health Volunteers (HV) and the MS cohort for each feature and at each difficulty levels. We will check how many features are statistically significant at p=0.005 for each features at each difficulty level. We use unpaired wilcoxon test because the same test is be done between HV and MS cohorts and this non-parametric test is being used to relax the normality assumption.


```{r  message=FALSE, warning=FALSE, include=FALSE}
## Construct a function to conduct the wilcoxon test between HV and MS cohorts
## and return features that are significant at each difficulty level

wtest_func <- function(data, alpha, group_var){
  ## Data is dataframe containing features and the first column being the categorical variables
  ## alpha is the alpha level (e.g 0.05)
  ## group_var is the grouping variable
  
  ## Remove irrelevent columns and rows
  sub_data <- data %>% 
    dplyr::select(difficulty_level, diagnosis_group, 
                  grep("F",names(data), value = TRUE)) %>% 
    dplyr::filter(diagnosis_group %in% c("HV","MS"))
  
  ## difficulty levels
  d_level <- c(1,2,3)
  ## empty vector of non-significant features
  DL1 <- c() 
  DL2 <- c() 
  DL3 <- c() 
  
  for (dl in d_level){
    sub_data.level <- sub_data %>% 
      dplyr::filter(difficulty_level==dl) %>% dplyr::select(-c(difficulty_level))
    f_numb <- dim(sub_data.level)[2]
    
    Group <- sub_data.level[,group_var]
    
    for (j in 2:f_numb){
      fi_data <- data.frame(Group = Group, Feature = sub_data.level[,j])
      res <- wilcox.test(Feature ~ Group, data = fi_data, exact = FALSE)
      pval <- res$p.value
      
      if (dl==1 & (pval < alpha)) {
        DL1 <- c(DL1, paste0("F", j-1, collapse = ""))
      }
      if (dl==2 & (pval < alpha)) {
        DL2 <- c(DL2, paste0("F", j-1, collapse = ""))
      }
      if (dl==3 & (pval < alpha)) {
        DL3 <- c(DL3, paste0("F", j-1, collapse = ""))
      }
    }
    
  }
  return(list(DL1 = DL1, DL2= DL2, DL3=DL3))
  
} 



```





```{r  message=FALSE, warning=FALSE, include=FALSE}
#### Wixcoxon test of difference between HV and MS
## Patient who use their dominant hand
sign.dom <- wtest_func(pat_feat.dom, 0.05, "diagnosis_group")
## Patient who did not use their dominant hand
sign.nondom <- wtest_func(pat_feat.nondom, 0.05, "diagnosis_group")



p_dom <- ggVennDiagram(sign.dom, show.legend = FALSE)
p_nondom <- ggVennDiagram(sign.nondom, show.legend = FALSE)

ggpubr::ggarrange(p_dom, p_nondom, 
          labels = c("Dominant", "Non-dominant"), ncol = 2, nrow = 1)

### Create dataframe of some features
pat_feat.dom.dl2 <- subset(pat_feat.dom, difficulty_level=="2" & diagnosis_group != "Others")
pat_feat.nondom.dl2 <- subset(pat_feat.nondom, difficulty_level=="2" & diagnosis_group != "Others")
dom_sub.ft <- data.frame(Category = rep("Dominant", dim(pat_feat.dom.dl2)[1]), 
                         Diagnosis = pat_feat.dom.dl2$diagnosis_group, 
                         Velocity = pat_feat.dom.dl2$F1, 
                         HausDError = pat_feat.dom.dl2$F24,
                         TTC = pat_feat.dom.dl2$F35)
nondom_sub.ft <- data.frame(Category = rep("Non-dominant", dim(pat_feat.nondom.dl2)[1]), 
                         Diagnosis = pat_feat.nondom.dl2$diagnosis_group, 
                         Velocity = pat_feat.nondom.dl2$F1, 
                         HausDError = pat_feat.nondom.dl2$F24,
                         TTC = pat_feat.nondom.dl2$F35)
sub.ft <- rbind(dom_sub.ft, nondom_sub.ft)

sub.ft.v <- sub.ft %>% dplyr::select(-c(HausDError, TTC))
sub.ft.h <- sub.ft %>% dplyr::select(-c(Velocity, TTC))
sub.ft.t <- sub.ft %>% dplyr::select(-c(HausDError, Velocity))

sub.ft.v2 <- data_summary(sub.ft.v, varname = "Velocity", groupnames = c("Category", "Diagnosis"))
sub.ft.h2 <- data_summary(sub.ft.h, varname = "HausDError", groupnames = c("Category", "Diagnosis"))
sub.ft.t2 <- data_summary(sub.ft.t, varname = "TTC", groupnames = c("Category", "Diagnosis"))


ggplot(sub.ft.v2, aes(x=Category, y=Velocity, fill=Diagnosis)) + 
  geom_bar(stat="identity", color="black", width = 0.5, position=position_dodge()) +
  geom_errorbar(aes(ymin=Velocity-sd, ymax=Velocity+sd), width=.2, position=position_dodge(.5)) +
  scale_fill_manual(values=c('blue','red')) +
  theme_bw() + Nice.Label + theme(legend.position="bottom")


ggplot(sub.ft.h2, aes(x=Category, y=HausDError, fill=Diagnosis)) + 
  geom_bar(stat="identity", color="black", position=position_dodge(), width = 0.5) +
  geom_errorbar(aes(ymin=HausDError-sd, ymax=HausDError+sd), width=.2, 
                position=position_dodge(.5)) + theme_bw() + Nice.Label +
  scale_fill_manual(values=c('blue','red')) +
  theme(legend.position="bottom")

ggplot(sub.ft.t2, aes(x=Category, y=TTC, fill=Diagnosis)) + 
  geom_bar(stat="identity", color="black", position=position_dodge(), width = 0.5) +
  geom_errorbar(aes(ymin=TTC-sd, ymax=TTC+sd), width=.2, position=position_dodge(.5)) +
  ylab("Time to Completion (s)") + scale_fill_manual(values=c('blue','red')) +
  theme_bw() + Nice.Label + theme(legend.position="bottom")
```




```{r  message=FALSE, warning=FALSE, include=FALSE}
## Obtain the list of all significant features
sign_features <- unique(c(sign.dom$DL1,sign.dom$DL2,sign.dom$DL3,
                          sign.nondom$DL1,sign.nondom$DL2,sign.nondom$DL3))
sign_features <- sign_features[order(as.numeric(gsub("F","", sign_features)))]

## Extract only significant features from patient and clinical features

pat_feat.dom.s <- pat_feat.dom %>% 
  dplyr::select(-setdiff(paste0("F",1:40,""), sign_features))
pat_feat.nondom.s <- pat_feat.nondom %>% 
  dplyr::select(-setdiff(paste0("F",1:40,""), sign_features))


```




## Correlation of Patient derived features

```{r  message=FALSE, warning=FALSE, include=FALSE}
## Function to extract matrix of features for correlation plot
mat.feat_func <- function(data, difficulty, diagnosis){
  ## dataframe containing features, and group category
  ## difficulty is difficulty level
  ## diagnosis is diagnosis group
  sub_data <- data %>% subset(difficulty_level==difficulty) %>% 
    dplyr::filter(diagnosis_group==diagnosis) %>% 
    dplyr::select(-c(1:10))
  return(as.matrix(sub_data))
} 

#### Extract HV and MS cohorts
## dominant hand
pat_feat.domHV_l1 <- mat.feat_func(pat_feat.dom.s,1,"HV")
pat_feat.domHV_l2 <- mat.feat_func(pat_feat.dom.s,2,"HV")
pat_feat.domHV_l3 <- mat.feat_func(pat_feat.dom.s,3,"HV")
pat_feat.domMS_l1 <- mat.feat_func(pat_feat.dom.s,1,"MS")
pat_feat.domMS_l2 <- mat.feat_func(pat_feat.dom.s,2,"MS")
pat_feat.domMS_l3 <- mat.feat_func(pat_feat.dom.s,3,"MS")
## non-dominant hand
pat_feat.nondomHV_l1 <- mat.feat_func(pat_feat.nondom.s,1,"HV")
pat_feat.nondomHV_l2 <- mat.feat_func(pat_feat.nondom.s,2,"HV")
pat_feat.nondomHV_l3 <- mat.feat_func(pat_feat.nondom.s,3,"HV")
pat_feat.nondomMS_l1 <- mat.feat_func(pat_feat.nondom.s,1,"MS")
pat_feat.nondomMS_l2 <- mat.feat_func(pat_feat.nondom.s,2,"MS")
pat_feat.nondomMS_l3 <- mat.feat_func(pat_feat.nondom.s,3,"MS")


```




```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}


#### Calculate the correlation matrix with pvalue
cor_type <- "spearman"
res.domHV_l1 <- rcorr(pat_feat.domHV_l1, type = cor_type)
res.domMS_l1 <- rcorr(pat_feat.domMS_l1, type = cor_type)

res.nondomHV_l1 <- rcorr(pat_feat.nondomHV_l1, type = cor_type)
res.nondomMS_l1 <- rcorr(pat_feat.nondomMS_l1, type = cor_type)


### Plot correlation plot of the features
col_cor<- colorRampPalette(c("red", "white", "blue"))(20)

corrplot(res.domHV_l1$r, type="upper", order="original", col = col,
         p.mat = res.domHV_l1$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2)

corrplot(res.domMS_l1$r, type="upper", order="original", col = col,
         p.mat = res.domMS_l1$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2)

corrplot(res.nondomHV_l1$r, type="upper", order="original", col = col,
         p.mat = res.nondomHV_l1$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2)

corrplot(res.domMS_l1$r, type="upper", order="original", col = col,
         p.mat = res.nondomMS_l1$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2)
 ### Plot selected features
my_dataHV <- pat_feat.nondomHV_l1[, c(1,25, 26)]
my_dataMS <- pat_feat.nondomMS_l1[, c(1,25, 26)]

PerformanceAnalytics::chart.Correlation(my_dataHV, histogram=TRUE, pch=19, method = "spearman")
PerformanceAnalytics::chart.Correlation(my_dataMS, histogram=TRUE, pch=19, method = "spearman")
```


## Statistical differences betweeen patient features at different difficulty levels


```{r  message=FALSE, warning=FALSE, include=FALSE}
## Construct a function to calculate statistical significant differences between 
## features at different levels given a dataframe

#### Create a function to plot a and b on top of the boxplot

## Function to generate letters based onm Tukey comparison test
tukey_label.func <- function(data ,TUKEY, group_var){
  ## Data is dataframe containing features and the first column being the categorical variables
  ## TUKEY is the Tukey test result
  ##group_var is the grouping variable in quotation mark
  
  # Extract labels and factor levels from Tukey post-hoc 
  Tukey.levels <- TUKEY[["Group"]][,4]
  Tukey.labels <- data.frame(multcompLetters(Tukey.levels)['Letters'])
  
  # Put the labels in the same order as in the data:
  Tukey.labels$treatment=rownames(Tukey.labels)
  Matching_Label<-data.frame(treatment=levels(data[,group_var]))
  Tukey.labels=Tukey.labels[match(Matching_Label$treatment,Tukey.labels$treatment),]
  #Tukey.labels=Tukey.labels[order(Tukey.labels$treatment), ]
  return(Tukey.labels)
}


## ANOVA test combined with Tuley comparison test
stat_diff.func <- function(data, alpha, group_var){
  ## Data is dataframe containing features and the first column being the categorical variables
  ## alpha is the alpha level (e.g 0.05)
  ## group_var is the grouping variable
  
  ## Extract the group variable
  Group <- data[,group_var]
  ## number of feature + grouping variable
  f_numb <- dim(data)[2] 
  
  ## Create an empty significance dataframe
  sign_df <- data[0,]
  sign_df[,group_var] <- as.character(sign_df[,group_var]) ## Convert group variable to character
  colnames(sign_df)[1] <- "" ## Remove first grouping variable name
  sign_df[1,1] <- "Significant"
  sign_df[2,1] <- "Level 1"
  sign_df[3,1] <- "Level 2"
  sign_df[4,1] <- "Level 3"
  
  
  for (j in 2:(f_numb-0)){
    ## Data frame of feature i and group
    fi_data <- data.frame(Group = Group, Feature = data[,j])
    
    ## Anova and kruskal wallis non-parametric test 
    anova_test <-aov(Feature~Group, data = fi_data)
    kw_test <- kruskal.test(Feature~Group, data = fi_data)
    
    ## Homogeneity of variance and normality of residual test from ANOVA
    levene_test <- LeveneTest(Feature~Group, data = fi_data) 
    shapiro_test <- shapiro.test(residuals(object = anova_test)) 
    
    ## Extract p-value from different tests
    pval.anova <- summary(anova_test)[[1]][["Pr(>F)"]][1]
    pval.kw <- kw_test$p.value
    pval.levene <- levene_test$`Pr(>F)`[1]
    pval.shapiro <- shapiro_test$p.value
    
    # is anova significant and assumptions satisfied
    if(pval.anova < alpha & pval.levene > alpha & pval.shapiro > alpha){ 
      ## Markm significant as yes
      sign_df[1,j] <- "YES"
      ## Tukey Multiple comparison test
      tukey_res = TukeyHSD(x=anova_test, 'Group', conf.level=0.95)
      ## Generate results from Tukey test
      label <- tukey_label.func(data = data,TUKEY= tukey_res,group_var)
      ## Add the letter crresponding to Tukey results
      sign_df[2:4,j] <- as.character(label$Letters) 
      # is anova non-significant and assumptions satisfied
    } else if (pval.anova >= alpha & pval.levene > alpha & pval.shapiro > alpha) {
      sign_df[1,j] <- "NO"
      sign_df[2:4,j] <- "-"
      # is one of the assumptions not satisfied but anova and kw significant
    } else if (pval.anova < alpha & (pval.levene < alpha | pval.shapiro < alpha) & pval.kw < alpha){
      sign_df[1,j] <- "YES"
      sign_df[2:4,j] <- as.character(label$Letters)
      # is one of the assumptions not satisfied and anova not sign. but kw significant
    } else if (pval.anova > alpha & (pval.levene < alpha | pval.shapiro < alpha) & pval.kw < alpha){
      sign_df[1,j] <- "YES (KW)"
       sign_df[2:4,j] <- "-"
      #anything else
    } else {
      sign_df[1,j] <- "NO"
      sign_df[2:4,j] <- "-"
    }
  }
  return(sign_df)
}

```



```{r  message=FALSE, warning=FALSE, include=FALSE}

# ## Cohort data with difficulty level
# pat_feat.HV_dl <- pat_feat.HV %>% 
#   dplyr::select(-c(patient_id,ntest_date,trial_id,gender,diagnosis_group,age))
# pat_feat.MS_dl <- pat_feat.MS %>% 
#   dplyr::select(-c(patient_id,ntest_date,trial_id,gender,diagnosis_group,age))
# pat_feat.Others_dl <- pat_feat.Others %>% 
#   dplyr::select(-c(patient_id,ntest_date,trial_id,gender,diagnosis_group,age))
# 
# ## Generate the dataframe of significance level based on difficulty level
# sign.df_HV <- stat_diff.func(pat_feat.HV_dl,0.05,"difficulty_level")
# 
# sign.df_MS <- stat_diff.func(pat_feat.MS_dl,0.05,"difficulty_level")
# 
# sign.df_Others <- stat_diff.func(pat_feat.Others_dl,0.05,"difficulty_level")
# 
# #### Significance based on diagnosis 
# pat_feat.diagnosis <- pat_feature %>% 
#   dplyr::select(-c(patient_id,ntest_date,difficulty_level,trial_id,gender,age))
# 
# ## significance at 0.05
# sign.df_diagnosis <- stat_diff.func(pat_feat.diagnosis,0.05,"diagnosis_group")

```



## Correlation of clinical derived features


```{r  message=FALSE, warning=FALSE, include=FALSE}
## Extract all clinical features



## Extract HV, MS, and Other cohorts
cl_feat.HV <- subset(cl_feature, diagnosis_group == "HV")
cl_feat.MS <- subset(cl_feature, diagnosis_group == "MS")
cl_feat.Others <- subset(cl_feature, diagnosis_group == "Others")

## Remove irrelevent columns
cl_feat.HV2 <- cl_feat.HV %>% dplyr::select(-c(1:10))
cl_feat.MS2 <- cl_feat.MS %>% dplyr::select(-c(1:10))
cl_feat.Others2 <- cl_feat.Others %>% dplyr::select(-c(1:10))

## significance at 0.05
# sign.df_diagnosis_cl <- stat_diff.func(cl_feature, 0.05, "diagnosis_group")

```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}


#### Calculate the correlation matrix with pvalue
res_cl.HV <- rcorr(as.matrix(cl_feat.HV2), type = cor_type)
res_cl.MS <- rcorr(as.matrix(cl_feat.MS2), type = cor_type)
res_cl.Others <- rcorr(as.matrix(cl_feat.Others2), type = cor_type)


corrplot(res_cl.HV$r, type="upper", order="original", col = col,
         p.mat = res_cl.HV$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2, na.label = "NA")

corrplot(res_cl.MS$r, type="upper", order="original", col = col,
         p.mat = res_cl.MS$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2, na.label = "NA")

corrplot(res_cl.Others$r, type="upper", order="original", col = col,
         p.mat = res_cl.Others$P, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2, na.label = "NA")

```


## Correlation between Patient derived features and clinical features


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}

## Function to extract and plot correplot between shape drawn feature and clinical features
mat.feat_func <- function(pat_data,cl_data, difficulty, diagnosis, dominant){
  ## pat_data is data of shape drawn feature and cl_data is clinical feature data
  ## difficulty is difficulty level
  ## diagnosis is diagnosis group
  ## dominant is if patient use dominant hand or not ("YES" or "NO")
  pat_sub <- pat_data %>% subset(difficulty_level==difficulty) %>% 
    dplyr::filter(dominant_hand_use==dominant, diagnosis_group==diagnosis) %>% 
    dplyr::select(-c(1:10))
  cl_sub <- cl_data %>% subset(difficulty_level==difficulty) %>% 
    dplyr::filter(dominant_hand_use==dominant, diagnosis_group==diagnosis) %>% 
    dplyr::select(-c(1:10))
  
  ## Calculate correlation matrix and corresponding p-value
  cor_mat <- rcorr(as.matrix(cl_sub), as.matrix(pat_sub), type = cor_type)
  r_pat_cl <- cor_mat$r[1:32,33:63]
  p_pat_cl <- cor_mat$P[1:32,33:63]
   ## plot
  corrplot(r_pat_cl, type="full", order="original", col = col,
         p.mat = p_pat_cl, sig.level = 0.05, insig = "blank", diag = TRUE,
         tl.col = "black", tl.cex = 1.1, font.lab=2, cl.cex = 1.2, na.label = "NA")
} 


############ plotting
#### Dominant Hand
## Healthy Cohorts
mat.feat_func(pat_feat.dom.s,cl_feature,1,"HV","YES")
mat.feat_func(pat_feat.dom.s,cl_feature,2,"HV","YES")
mat.feat_func(pat_feat.dom.s,cl_feature,3,"HV","YES")
## MS Cohorts
mat.feat_func(pat_feat.dom.s,cl_feature,1,"MS","YES")
mat.feat_func(pat_feat.dom.s,cl_feature,2,"MS","YES")
mat.feat_func(pat_feat.dom.s,cl_feature,3,"MS","YES")

#### Non-dominant Hand
## Healthy Cohorts
mat.feat_func(pat_feat.nondom.s,cl_feature,1,"HV","NO")
mat.feat_func(pat_feat.nondom.s,cl_feature,2,"HV","NO")
mat.feat_func(pat_feat.nondom.s,cl_feature,3,"HV","NO")
## MS Cohorts
mat.feat_func(pat_feat.nondom.s,cl_feature,1,"MS","NO")
mat.feat_func(pat_feat.nondom.s,cl_feature,2,"MS","NO")
mat.feat_func(pat_feat.nondom.s,cl_feature,3,"MS","NO")

```


