# Introduction

Smartphone devices may be an easier alternative to obtain data from patients outside of clinics and hospitals environment. In this project, smartphone-based data were obtained from Multiple Sclerosis (MS) patients using a drawing a spiral test. The test was administered in order to measure upper extremity functionality. 

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
### Libraries
library(openxlsx) ## v. 4.1.5
library(dplyr) ## v. 1.0.0
library(ggplot2) ## v. 3.3.1
library(wavelets) ## v. 0.3.0.2 ## for wavelet transform -- https://cran.r-project.org/web/packages/wavelets/wavelets.pdf
library(signal) ## v. 0.7.6 ## for filtering -- https://cran.r-project.org/web/packages/signal/signal.pdf
library(e1071) ## v. 1.7.3
library(psd) ## v. 2.1.0 ## power spectral density -- https://cran.r-project.org/web/packages/psd/psd.pdf
library(TSEntropies) ## v. 0.9 ## to calculate Approximate Entropies (ApEn)-- https://cran.r-project.org/web/packages/TSEntropies/TSEntropies.pdf
library(oce) ## v.1.2.0 ## oceanographic library which contain PSD with hamming window
library(tidyr) ## v. 1.1.0 ## for gather and wide
library(VennDiagram) ## v.1.6.20 ## Venn diagram 
library(forecast) ## v. 8.13 ## for outlier detection https://cran.r-project.org/web/packages/forecast/forecast.pdf
library(fda.usc) ## v. 2.0.2
library(pathmapping) ## v. 1.0.2 ## For Area Under two curves ##https://cran.r-project.org/web/packages/pathmapping/pathmapping.pdf
library(DescTools) ## v. 0.99.38 ## ## For trapezoid rule https://www.rdocumentation.org/packages/DescTools/versions/0.99.37
#library(SpatEntropy) ##v.0.1.0 ## For Image Entropy ## https://cran.r-project.org/web/packages/SpatEntropy/SpatEntropy.pdf
#library(raster) ## v. 3.3.13 ## to create a grid of x y, coordinates of z values
library(corrplot) ##v.0.84 ## for correlation matrix plot
library(Hmisc) ## v.4.4.0 ## to calculate matrix of correlation and pvalue
library(multcompView) ## v. 0.1.8 ## to generate letters from Tukey test
library(CulturalAnalytics) ## v. 1.0.7 ## for image entropy
library(entropy) ## v. 1.2.1 ## To use discretize2d for 2D binning 
library(SpatEntropy) ##v. 0.1.0 ## For Image entropy using shannon
library(ggVennDiagram) ## v. 0.3 ## for ggplot vendiagram
```


## Data cleaning

First, we cleaned up the data by correcting some of the patient ids that were entrered incorrectly. For example, NIB380 on 5/22/19  was changed to NIB370, NIB649 on 3/14/19 was changed change to NIB640, NIB685 on 12/12/18 was changed to NDS68, and NIB703 on 8/21/19 was changed to NDS703. Also, all extreme coordinate points that were extremly outside of the real spiral were removed (i.e. y < -0.8 and abs(x) > 1.4 ). All patients with name starting with ``PATIENT" who are healthy volunteers were extracted from the data since they do not have clinical informations. The clinical data and the spiral multiple sclerosis were then merged. 


```{r  message=FALSE, warning=FALSE, include=FALSE}
### Import data
## Description of the data is as follow
# Patient = Subject ID
# x: distance (in inches) of the spot drawn from the center of the screen
# y: distance (in inches) of the spot drawn from the center of the screen
# p: estimated pressure of the tap (based off of surface area)
# t: UNIX timestamp of when the drawing happened (to the millisecond)
# sumData.Num turns: the number of turns in the spiral 
# sumData.Line width: the width of the lines of the spiral
# time.point: UNIX timestamp of the drawing (this time in seconds)



ms_data <- read.csv("spiralDataCorrected.csv", header = TRUE) ## multiple sclerosis
cl_data <- read.csv("Clinical_Data.csv", header = TRUE) ## clinical


#### Clean up the MS data
## add a test date to the data 
ms_data$testdate <- substr(ms_data$testDate,1,10)

## Remove non-relevant rows that have patient ID such as x:1234455
ms_data <- ms_data[substr(ms_data$patientID,1,2) != "x:" & 
                     substr(ms_data$patientID,1,2) != "{t" & 
                     substr(ms_data$patientID,1,2) != "y:" &
                     substr(ms_data$patientID,1,2) != "z:" &
                     substr(ms_data$patientID,1,2) != "t:" &
                     substr(ms_data$patientID,1,2) != "" &
                     substr(ms_data$patientID,1,11) != "UPLOAD_DATE" &
                     substr(ms_data$patientID,1,9) != "Altitude:",]

## and convert variable to numeric
ms_data$x <- as.numeric(as.character(ms_data$x))
ms_data$y <- as.numeric(as.character(ms_data$y))
ms_data$t <- as.numeric(as.character(ms_data$t))
ms_data$p <- as.numeric(as.character(ms_data$p))
ms_data$age <- as.numeric(as.character(ms_data$age))
ms_data$testdate <- as.Date(ms_data$testdate, format = "%Y-%m-%d")

### Certain patients are incorrect so they are changed
### NIB380 on 5/22/19 – change the patient code to NIB370 
### NIB649 on 3/14/19 – change the patient code to NIB640 
### NIB685 on 12/12/18 – change the patient code to NDS685 
### NIB703 on 8/21/19 – change the patient code to NDS703 

ms_data[ms_data$patientID =="NIB380" & ms_data$testdate=="2019-05-22",]$patientID <- "NIB370"
ms_data[ms_data$patientID =="NIB649" & ms_data$testdate=="2019-03-14",]$patientID <- "NIB640"
ms_data[ms_data$patientID =="NIB685" & ms_data$testdate=="2018-12-12",]$patientID <- "NDS685"
ms_data[ms_data$patientID =="NIB703" & ms_data$testdate=="2019-08-21",]$patientID <- "NDS703"

## Retrive some of the healthy volunteers that start with PATIENT..
HV_Patient <- ms_data[substr(ms_data$patientID,1,7)=="PATIENT",]

```




```{r  message=FALSE, warning=FALSE, include=FALSE}
## add visit number based on patientID and difficulty level, remove some columns and change column names
ms_data2 <- ms_data %>% 
  group_by(patientID, difficulty) %>%
  mutate(ntest_date = factor(testdate, labels = 1:length(unique(testdate)))) %>% 
  dplyr::select(trialID, patientID,x,y,t,p,ntest_date,difficulty, testdate, appendage) %>% 
  dplyr::rename(trial_id = trialID, patient_id=patientID,difficulty_level=difficulty)

ms_data2$t <- ms_data2$t/1000 ## convert time point from milliseconds to seconds 

## add time that is scaled from 0 by patient_id, ntest_date, and difficulty_level and remove duplicate x, y, t
ms_data2 <- ms_data2 %>% 
  group_by(patient_id, ntest_date, difficulty_level, trial_id) %>% 
  distinct(x,y,t,p, .keep_all = TRUE) %>% 
  arrange(t) %>% 
  mutate(pixel = row_number(), time = t-t[1])

```




```{r  message=FALSE, warning=FALSE, include=FALSE}

####### Clinical data

#### Extrract the clinical label names and change them to CL1
cl_dat <- cl_data
cl_labels <- colnames(cl_dat)
cl_labels.sub <- cl_labels[-c(1:9,40)]

## new clinical labels
new_cl_label <- paste(rep("CL",length(cl_labels.sub)),1:length(cl_labels.sub),sep = "")

## Create index and replace the cl labels with index 
new_cl_index <- match(cl_labels.sub, cl_labels)
colnames(cl_data)[new_cl_index] <- new_cl_label

## To recognize which label is what column name
cl_label_table <- data.frame(Label = new_cl_label, Name = cl_labels.sub)

#### Add new diagnosis category to the clinical data 
#### (MS, HV, and Others for all others disorder)
cl_data$diagnosis_group <- ifelse(
  cl_data$diagnosis=="PP-MS" | cl_data$diagnosis=="RR-MS" | cl_data$diagnosis=="SP-MS","MS",
                                   ifelse(cl_data$diagnosis=="Healthy Donor","HV", "Others"))

## round age to the nearest integer
cl_data$age <- round(cl_data$age)

## Create clinical data that contain patient that intersect with those in the ms_data
## CL6=X9HPT.LH.Avg + CL7=X9HPT.RH.Avg
cl_data.sub <- cl_data %>% 
  dplyr::rename(dominant_hand = handedness, patient_id=NeurExPatientID, 
                testdate = "AppTestDate", neuro_testdate ="NeurExDate",
                apptest_to_neurotest = "X..days.between.app.test.and.clinic.visit") %>% 
  mutate(X9HPT.Avg = (CL6+CL7)/2)


## Extracting the MS patients from clinicalm dataset
cl_data_patient <- subset(cl_data.sub, diagnosis_group =="MS") 
cl_data_patient <- mutate(cl_data_patient, id = row_number())

```



```{r  message=FALSE, warning=FALSE, include=FALSE}

### Splitting the dataset 2/3 training and 1/3 test sets
set.seed(42)

## Create Training set (train = 2/3)
cl_train <- cl_data_patient %>% slice_sample(prop = 0.67, weight_by = X9HPT.Avg)

## Create test set (test = 1/3)
cl_test  <- anti_join(cl_data_patient, cl_train, by = 'id')


######
## Create a function to merge clinical and MS datasets
merge_func <- function(cl_ms, cl_data, ms_data){
  ## cl_ms is the patient MS datatset (train or test set)
  ## cl_data is the entire clinical dataset
  ## ms_data is the MS dataset (i.e. spiral dataset)

  ## healthy volunters and other disorders clinical data
  cl_hv_other <- subset(cl_data, diagnosis_group=="HV" | diagnosis_group=="Others")
  
  ## Combine clinical MS patient with other disease and healthy volunteers
  cl_ms <- cl_ms %>% dplyr::select(-c(id)) ## remove id column before rbind
  cl_ms_hv_other <- rbind(cl_ms, cl_hv_other)
  
  ## Change the format of the test dates and check difference between test dates (app and neuro)
  cl_ms_hv_other$testdate <- as.Date(cl_ms_hv_other$testdate, format = "%m/%d/%Y")
  cl_ms_hv_other$neuro_testdate <- as.Date(cl_ms_hv_other$neuro_testdate, format = "%m/%d/%Y")
  
  ## patient and testdates in both datasets
  patient_ms_cl <- intersect(ms_data$patient_id,cl_ms_hv_other$patient_id) 
  testdate_ms_cl <- intersect(ms_data$testdate,cl_ms_hv_other$testdate) 
  
  ## Clinical datasets with patient IDs and testdates in MS datasets
  cl_ms_hv_other <- cl_ms_hv_other %>% 
    dplyr::filter(patient_id %in% patient_ms_cl & testdate %in% testdate_ms_cl) 

  ## MS datasets with patient IDs and testdates in clinicla datasets
  ms_data <- ms_data %>% semi_join(cl_ms_hv_other, by = c("patient_id","testdate"))

  #### Join the cl_data.sub to the ms_data2 
  ms_data_merge <- merge(ms_data, cl_ms_hv_other, by = c("patient_id", "testdate"), 
                         all = TRUE)

  ## Arrange ms_data3 by patient_id,ntest_date, difficulty_level, trial_id, and time
  ms_data_merge <- arrange(ms_data_merge,patient_id,ntest_date, 
                    difficulty_level, trial_id, time)
  
  return(ms_data_merge)
}


## Create Training set (train = 2/3)
ms_data_train <- merge_func(cl_ms = cl_train, cl_data = cl_data.sub, ms_data = ms_data2)

## Create test set (test = 1/3)
ms_data_test <- merge_func(cl_ms = cl_test, cl_data = cl_data.sub, ms_data = ms_data2)

##### Remove lines with y coordinate < -0.8 and x coordinates > 1.2 so completely
#### outside of the the orange real spiral (i.e. it is an error)

```

## Removing outliers
In what follow we removed outliers using the y coordinates points by first identifying residuals from loess fitting curve. Then the residuals are labelled as outliers if they lie outside of the range $\pm 2(q_{0.9}- q_{0.1}$ where $q_p$ is the p quantile of the residuals. The procedure was done using the ``tsoutliers" library and more can be seen regarding that in [tsoutliers documentation](https://robjhyndman.com/hyndsight/forecast5/#:~:text=A%20more%20general%20function%20is,values%20replaced%20by%20estimated%20values). 


```{r  message=FALSE, warning=FALSE, include=FALSE}

## Function to detect and remove outlier from a given trial
outlier_func <- function(dataframe){
  ## dataframe is a data frame containing y coordinate
  data_with_id <- dataframe %>% dplyr::mutate(id = row_number())
  outlier_index <- tsoutliers(data_with_id$y, lambda = "auto")$index
  
  if (length(outlier_index) == 0){
    new_data = data_with_id %>% dplyr::select(-id)
  } else {
    new_data = data_with_id %>% dplyr::filter(id != outlier_index) %>% dplyr::select(-id)
  }
  return(new_data)
}


## Remove the outlier by trial group 
ms_data3 <- ms_data_train %>% 
  group_by(patient_id, ntest_date, difficulty_level, trial_id) %>%
  group_modify(~outlier_func(.x))

```




```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}
#### create template for plot labelling
black.bold.text1 <- element_text(face = "bold", color = "black",size=24) # x and y axis
black.bold.text2 <- element_text(face = "bold", color = "black",size=24) # title
Nice.Label <-theme(axis.text.x = element_text(face="bold", color="black", size=16),
         axis.text.y = element_text(face="bold", color="black", size=16),
         title=black.bold.text2,axis.title = black.bold.text1, legend.position = "right",legend.box = "vertical",
         legend.text = element_text(size=24),strip.text.x = element_text(face="bold",size=24)) #18 or 24

### Tranform from long tabel to wide table

metadata_sub.long <- tibble::as_tibble(ms_data3) %>% 
  dplyr::select(patient_id,diagnosis_group,dominant_hand, gender) %>%
  dplyr::distinct(patient_id, .keep_all = TRUE) %>%
  tidyr::gather(group, attribute, diagnosis_group:gender, factor_key = TRUE) 

metadata_sub.long$group <- factor(metadata_sub.long$group,
                                  levels = c("diagnosis_group","gender","dominant_hand"),
                                  labels = c("Diagnosis Categories", "Gender", "Dominant Hand"))

### Plot the distribution of the type of data we have 
ggplot(metadata_sub.long, aes(x=attribute)) +
  geom_bar(colour = "gray", width = 0.5) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.2, size = 6) +
  facet_wrap(~group, scales = "free", nrow = 1) +
  labs(x = " ", y = "Number of Cohorts") +
  theme_bw()+Nice.Label

```



## True spiral data from the App

The true spiral data from the app was generated using a function that take two parameters as input: a Side (Left or Right) and a difficulty level (in the set of 1, 2, 3).  The side is turned into the invertFactor, and the difficulty is turned into the number of turns the spiral will make, either 2, 3, or 5 (as well as the line thickness).  AngleToRadiusScale is then derived from the maxRadius and the maxAngleRadians, which is derived from the number of turns.  Then to generate the points, the function loops over all the possible angle values in steps of .1 radians and does the coordinate math using the following formula (see [wikipedia](https://en.wikipedia.org/wiki/Polar_coordinate_system#Converting_between_polar_and_Cartesian_coordinates) for more explanation):
$$ x =  r \cos{\phi} \\ y = r\sin{\phi}$$
Inside this loop is where the "current radian value" is used. Thus, to generate the spiral the only parameters needed are Side and Difficulty, combined with the constants of angleOffset, maxRadius, and angle increment (.1 radians). 

```{r  message=FALSE, warning=FALSE, include=FALSE}

#### Recreate the original spiral drawing from the App

## Turn Difficulty level into Number of Turns
DifficultyToNumberOfTurns <- function(difficulty){
  if(difficulty=="3"){
    return(5)
  } else if (difficulty == "2"){
    return(3)
  } else {
    return(2)
  }
}

## Turn difficulty level into the Line width
DifficultyToLineWidth <- function(difficulty){
  if(difficulty=="1"){
    return(6/32)
  } else if (difficulty == "2"){
    return(5/32)
  } else {
    return(4/32)
  }
}


## Function to generate the spiral data
GenerateSpiral = function(side, difficulty){
  ## side can either be LH or RH from the appendage in ms_data
  ## difficulty is the dificulty level from 1,2 or 3
  
  NumberOfTurns <- DifficultyToNumberOfTurns(difficulty)
  MaxAngleRadians <- 2*pi*NumberOfTurns
  MaxRadius <- 1.25
  AngleToRadiusScale <- MaxRadius/MaxAngleRadians
  AngleOffset <- pi/2
  InvertFactor <- ifelse(side=="LH",1,-1)
  
  CurAngleRadian <- 0
  x_vec <- vector() ## Initialize the x vector
  y_vec <- vector() ## initialize the y vector
  
  while (CurAngleRadian < MaxAngleRadians) {
    x <- CurAngleRadian*AngleToRadiusScale*cos((CurAngleRadian * InvertFactor) + AngleOffset)
    y <- CurAngleRadian*AngleToRadiusScale*sin((CurAngleRadian * InvertFactor) + AngleOffset)
    
    x_vec <- append(x_vec,x)
    y_vec <- append(y_vec,y)
    CurAngleRadian = CurAngleRadian + 0.1
  }
  
  linewidth <- rep(DifficultyToLineWidth(difficulty), length(x_vec))
  
  return(list(x=x_vec, y=y_vec, linewidth = linewidth))
}


```


## Scaling of the X and Y coordinates

In what follow, we scaled the x and y coordinates according to the true spiral data. We assumed that every first point of the patient in the attempt to draw is the spiral is correct and thus the first point is shifted toward the first point of the true spiral data. Every other point has been shifted accordingly.


```{r  message=FALSE, warning=FALSE, include=FALSE}
#### Create a function that remove all errors and normalize x and y coordinates

normXY_func <- function(data){
  ## data is dataframe containing the x and y coordinates
  
  ## factor levels of patient ID
  p.id <- levels(factor(data$patient_id))
  
  # create an empty dataframe to be used later
  new.tr.level <- data[0,]
  
  #### Initialize the scaled x and y coordinates
  ## Based on first points of the patient
  data$x_scf <- rep(0, dim(data)[1])
  data$y_scf <- rep(0, dim(data)[1])
  ## Shift to center for all spiral
  data$x_scc <- rep(0, dim(data)[1])
  data$y_scc <- rep(0, dim(data)[1])
  
  for (id in p.id) {
    patient_data <- subset(data, patient_id==id) ## select a single patient
    n.test <- levels(factor(patient_data$ntest_date))
    for (tn in n.test) {
      test_date <- subset(patient_data, ntest_date==tn) ## select one test date
      d.levels <- levels(factor(test_date$difficulty_level))
      for (dl in d.levels) {
        d.level <- subset(test_date, difficulty_level==dl) ## select difficulty level
        tr.levels <- levels(factor(d.level$trial_id))
        for (tr in tr.levels){
          tr.level <- subset(d.level, trial_id==tr) ## select trial_id
          
          dim_tr <- dim(tr.level)[1] ## length of the trial level
          
          appendage <- tr.level$appendage[1]
          spiral_dl <- as.data.frame(GenerateSpiral(appendage,dl))
          
          #### use this if first point is outward
          # shitf based on first point
          norm.xo <- last(spiral_dl$x) - tr.level$x[1]
          norm.yo <- last(spiral_dl$y) - tr.level$y[1]
          # shift based on center point
          norm.xo2 <- first(spiral_dl$x) - tr.level$x[dim_tr]
          norm.yo2 <- first(spiral_dl$y) - tr.level$y[dim_tr]
          
          #### use this if first point is inward
          norm.xi <- first(spiral_dl$x) - tr.level$x[1]
          norm.yi <- first(spiral_dl$y) - tr.level$y[1]
          
          ## scale the x and y coordinates
          
          for (j in 1:(dim_tr)) {
            if (tr.level$y[1] > tr.level$y[dim_tr]){ ## outward loop
              tr.level[j, "x_scf"] <- tr.level$x[j] + norm.xo
              tr.level[j, "y_scf"] <- tr.level$y[j] + norm.yo
              tr.level[j, "x_scc"] <- tr.level$x[j] + norm.xo2
              tr.level[j, "y_scc"] <- tr.level$y[j] + norm.yo2
            } else if (tr.level$y[1] < tr.level$y[dim_tr]){ ## inward loop
              tr.level[j, "x_scf"] <- tr.level$x[j] + norm.xi
              tr.level[j, "y_scf"] <- tr.level$y[j] + norm.yi
              tr.level[j, "x_scc"] <- tr.level$x[j] + norm.xi
              tr.level[j, "y_scc"] <- tr.level$y[j] + norm.yi
            } else { ## anything else
              tr.level[j,"x_scf"] <- NA
              tr.level[j,"y_scf"] <- NA
              tr.level[j,"x_scc"] <- NA
              tr.level[j,"y_scc"] <- NA
            }
            
          }
          new.tr.level <- rbind(new.tr.level, tr.level) ## append the data
        }
      }
    }
  }
  
  return(as.data.frame(new.tr.level))
  
}



####
# ptm <- proc.time() ## Check how long it will run (takes 38.8 min to run)
# 
# ms_data.xy_sc <- normXY_func(ms_data3)
# 
# proc.time() - ptm
# # 
# 
# ## Save to be use later
# write.csv(ms_data.xy_sc, file = "data_xy_scale.csv")

  
```





