# Spatial features extraction

## Hausdorff Distance (dH(X,Y))
We computed drawing error based on a shape-matching approach known as the Hausdorff distance.First, let $X$ and $Y$ be two-non empty subsets of a metric space $(M, d)$ where $M$ is set and $d$ is a metric or distance on $M$. The formula of the Hausdorff distance $d_H(X,Y)$ is as follow:
$$d_H(X,Y) = \max\{\sup_{x\in X} \inf_{y\in Y} d(x,y), \sup_{y\in Y}\inf_{x\in X} d(x,y)\}$$
From Creagh et al., 2020, it was stated that ``This metric compares the maximum distance of one set to the nearest point in another set, which can be used as a basis to compute the error between the reference way-points (interpolated into a reference shape scaled to the number of pixels drawn) and the subject’s drawing attempt". Thus wqe took similar approach by first interpolating the reference points to the size of the drawing attempt prior to calculating the $d_H(X,Y)$.

```{r  message=FALSE, warning=FALSE, include=FALSE}

## Function to set up the data frame to later calculate the Hausdorff distances
HausD_dataframe <- function(data, xs, ys){
  ## data is the dataframe containing drawn x and y coordinates
  ## xs and ys are the x and y coordinates
  
  appendage <- data$appendage[1]
  dl <- data$difficulty_level[1]
  spiral_dl <- GenerateSpiral(appendage,dl)
  
  ## interpolate the reference spiral to the size of x coordinate
  # Drawn coordinates
  x <- data[, xs]
  y <- data[, ys]
  dat <- fdata(data.frame(x1=x,y1=y)) ## convert to functional data class
  # reference coordinate interpolated to the lengh of x
  ref_x <- spline(spiral_dl$x, n=length(x))$y
  ref_y <- spline(spiral_dl$y, n=length(y))$y
  ref_dat <- fdata(data.frame(x2=ref_x,y2=ref_y)) ## convert to functional data class
  
  
  return(list(drawing_df = dat, ref_df = ref_dat))
}

####

HausData_func <- function(data, xs, ys){
  ## data is the entire dataset with x and y scaled
  ## xs is the x coordinates to use (either x_scf or x_scc)
  ## ys is the y coordinates to use (either y_scf or y_scc)
  
  #### intialize an empty 0 column
  data[,"hausD"] <- 0 ##  Maximum H_d
  data[,"hausDError"] <- 0 ## Sum of  H_d (error)
  data[, "hausDt"] <- 0 ## H_d normalized by time taking to complete the drawing
  data[,"hausDIqr"] <- 0 ## Interquartile range of H_d
  data[, "hausD25"] <- 0 ## Beginning error (25%)
  data[, "hausD75"] <- 0 ## Ending error (75%)
  data[,"hausDmiddle"] <- 0 ## H_d from the middle (15% to 85%)
  data[,"hausDmiddlet"] <- 0 ## Middle H_d(15% to 85%) weighted by time to completion

  
  
  # create an empty dataframe to be used later
  new.tr.level <- data[0,]
  
  new.tr.level <- new.tr.level %>% 
    dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i))
  
  ## factor levels of patient ID
  p.id <- levels(factor(data$patient_id))
  
  for (id in p.id) {
    patient_data <- subset(data, patient_id==id) ## select a single patient
    n.test <- levels(factor(patient_data$ntest_date))
    for (tn in n.test) {
      test_date <- subset(patient_data, ntest_date==tn) ## select one test date
      d.levels <- levels(factor(test_date$difficulty_level))
      for (dl in d.levels) {
        d.level <- subset(test_date, difficulty_level==dl) ## select difficulty level
        tr.levels <- levels(factor(d.level$trial_id))
        #ct <- 0 ## Intialize count to be used in the trial level
        for (tr in tr.levels){
          tr.level <- subset(d.level, trial_id==tr) ## select trial_id
          dim_tr <- dim(tr.level)[1] ## Make sure the dimension of each trial > 2
          if (dim_tr>2){
            
            #ct = ct + 1 ## start a count for increment
           #print(ct) ## to check the loop is working fine 
            
            ## Get the drawing and reference data
            dat <- HausD_dataframe(tr.level, xs, ys)$drawing_df$data
            ref_dat <- HausD_dataframe(tr.level, xs, ys)$ref_df$data
            ## Get the total time use to complete th drawing
            time_n <- tr.level$time[dim_tr]
            
            ## Subset data from begining 25%, ending 75%, and middle 15%-85%
            dat25 <- fdata(dat[1:round(dim_tr*0.25),])
            dat75 <- fdata(dat[round(dim_tr*0.75):dim_tr,])
            dat15_85 <- fdata(dat[round(dim_tr*0.15):round(dim_tr*0.85),])
            
            ref_dat25 <- fdata(ref_dat[1:round(dim_tr*0.25),])
            ref_dat75 <- fdata(ref_dat[round(dim_tr*0.75):dim_tr,])
            ref_dat15_85 <- fdata(ref_dat[round(dim_tr*0.15):round(dim_tr*0.85),])
            
            
            #### Calculate the different metrics
            tr.level[tr.level$trial_id==tr,"hausD"] <- 
              max(metric.hausdorff(fdata(dat),fdata(ref_dat)), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDError"] <- 
              sum(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDt"] <-
              max(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)/time_n
            tr.level[tr.level$trial_id==tr,"hausDIqr"] <-
              IQR(metric.hausdorff(fdata(dat),fdata(ref_dat)),na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausD25"] <- 
              max(metric.hausdorff(dat25,ref_dat25), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausD75"] <- 
              max(metric.hausdorff(dat75,ref_dat75), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDmiddle"] <-
              max(metric.hausdorff(dat15_85,ref_dat15_85), na.rm = TRUE)
            tr.level[tr.level$trial_id==tr,"hausDmiddlet"] <-
              max(metric.hausdorff(dat15_85,ref_dat15_85), na.rm = TRUE)/time_n
            
            ## Remove some variables and take first row only
            tr.level <- tibble::as_tibble(tr.level) %>% 
              dplyr::select(-c(t,time,x,y,x_scf,y_scf,x_scc,y_scc,d_t,v_i,rv_i,av_i)) %>%
              group_by(patient_id, ntest_date, difficulty_level, trial_id) %>% 
              slice(1)
            
            tr.level <- as.data.frame(tr.level) ## convert to data.frame
            
            new.tr.level <- rbind(new.tr.level, tr.level) ## append the data
             
          }
        }
      }
    }
  }
  
  return(as.data.frame(new.tr.level))  
}


```









```{r  message=FALSE, warning=FALSE, include=FALSE}

#### calculate data frame with hausdorff distance metrics

## Coordinates x and y
ptm <- proc.time() ## Take 56.85 min
ms_data.hausD_xy <- HausData_func(ms_data4, "x", "y")
proc.time() - ptm
#
write.csv(ms_data.hausD_xy, file = "ms_data.hausD_xy.csv")


## Coordinates x_scf and y_scf
ptm <- proc.time() ## Take 59 min
ms_data.hausD_xscf <- HausData_func(ms_data4, "x_scf", "y_scf")
proc.time() - ptm


write.csv(ms_data.hausD_xscf, file = "ms_data.hausD_xscf.csv")


## Coordinates x_scc and y_scc
ptm <- proc.time() ## Take 59 min
ms_data.hausD_xscc <- HausData_func(ms_data4, "x_scc", "y_scc")
proc.time() - ptm
# 
# 
write.csv(ms_data.hausD_xscc, file = "ms_data.hausD_xscc.csv")

```




## Approximate Entropy (ApEn)


```{r  message=FALSE, warning=FALSE, include=FALSE}

### Approximate Entropy (ApEn)
## Interpolate all v_i, r_vi, av_i over a fixed length L = 500 (see similar approach in Creagh et al., 2020) and calculate their entropy 
ms_data.apen <- ms_data4 %>% 
   group_by(patient_id,ntest_date,difficulty_level, trial_id) %>%
  summarise(v_apen = ApEn(spline(v_i, n = 500)$y), 
            rv_apen = ApEn(spline(rv_i, n = 500)$y), 
            av_apen = ApEn(spline(av_i, n = 500)$y))


p1.apen <- subset(ms_data.apen, patient_id=="NIB632" & ntest_date=="1")
p1.apen <- p1.apen %>% select(patient_id, difficulty_level, v_apen)
p1.apen$difficulty_level <- as.factor(p1.apen$difficulty_level)
  
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

####

p1.apen.sum <- data_summary(p1.apen, "v_apen", "difficulty_level")
  
ggplot(p1.apen.sum, aes(x=difficulty_level, y=v_apen)) + 
    geom_bar(stat="identity", color="black", colour = "gray", width = 0.5, 
           position=position_dodge()) +
  labs(x = "Difficulty Levels", y = "Approximate Speed Entropy") +
    geom_errorbar(aes(ymin=v_apen-sd, ymax=v_apen+sd), width=.2, size = 0.9,
                  position=position_dodge(0.05))+ theme_bw() + Nice.Label


###

```

