# Modification of Features

In what follow, both clinical and patient derive features will be modify to create other features for analysis previously conducted

## Selecting appropriate clinical features

```{r  message=FALSE, warning=FALSE, include=FALSE}

## Create new clinical features as combinations of old features and rearrange features
new_cl.HV.MS <- cl_feature.HV.MS %>% 
  dplyr::select(patient_id, testdate, diagnosis_group, dominant_hand, C1, C8, C9, C6, C7,
                C22,C5, C10, C3, C12, C13, C14, C15, C4, C16, C17, C20, C21, C23,
                C24, C25, C26, C27, C28) %>% 
  dplyr::mutate(nC10 = C12 + C14, nC11 = C13 + C15, 
                testdate = as.Date(testdate, format = "%m/%d/%Y")) %>% 
  dplyr::rename(oC3=C3, oC4=C4, oC12=C12, oC13=C13, oC14=C14, oC15=C15) %>% 
  dplyr::rename(C2=C8, C3=C9,C4=C6, C5=C7,C6=C22, C7=C5, C8=C10, C9=oC3, C10=nC10, 
                C11=nC11, C12= oC4, C13=C16, C14=C17, C15=C20, C16=C21,
                MRI1=C23, MRI2=C24, MRI3=C25, MRI4=C26, MRI5=C27, MRI6=C28) %>% 
  dplyr::select(patient_id, testdate, diagnosis_group, dominant_hand, C1, C2, C3, C4, 
                C5, C6, C7, C8, C9, C10, C11, C12, C13, C14, C15, C16,
                MRI1, MRI2, MRI3, MRI4, MRI5, MRI6)
  
##### Statistical Significant Features
wtest_func_onedl(new_cl.HV.MS, 0.001, "diagnosis_group",paste(c("C","MRI"), collapse = "|"))

### Remove features that are not significant
new_cl.HV.MS.S <- new_cl.HV.MS %>% 
  dplyr::select(-c(C8, MRI1, MRI2, MRI3))

```

## Calculating additional patient-derived features

```{r  message=FALSE, warning=FALSE, include=FALSE}
## function to scale between 0 and 1
scale01 <- function(x){return((x-min(x, na.rm = TRUE))/(max(x, na.rm = TRUE)-min(x, na.rm = TRUE)))} 

### Calculate the scale sum of Hausdorff + scaled time taken to complete drawings
newfull_feature_train <- full_feature_train %>% 
  dplyr::group_by(patient_id, difficulty_level, diagnosis_group, dominant_hand_use) #%>% 
  #dplyr::mutate(F41 = log(F24 + 1, base = 2), F42 = scale01(F24) + scale01(F35))

newfull_feature_test <- full_feature_test2 %>%  
  dplyr::group_by(patient_id, difficulty_level, diagnosis_group, dominant_hand_use) #%>% 
  #dplyr::mutate(F41 = log(F24 + 1, base = 2), F42 = scale01(F24) + scale01(F35)) 

```

## Calculate the interclass correlation Coefficient (ICC)


```{r  message=FALSE, warning=FALSE, include=FALSE}
## Extract HV and MS patients with longitudinal data
nlongi_pat <- newfull_feature_train
nlongi_pat$ntest_date <- as.numeric(newfull_feature_train$ntest_date)
nlongi_pat <- unique(subset(nlongi_pat, ntest_date >= 6)$patient_id)

nlongi_data <- subset(newfull_feature_train, patient_id %in% nlongi_pat)

### Average features by trial_id group
nlongi_data.av <- nlongi_data %>% 
  dplyr::select(-c(trial_id, age, gender, appendage, dominant_hand)) %>% 
  group_by(patient_id, ntest_date, difficulty_level, diagnosis_group, dominant_hand_use) %>% 
  dplyr::summarise_all(.funs = mean)


##Extract HV and MS   
nlongi_data.HV <- subset(nlongi_data.av, diagnosis_group=="HV")


## Extract dominant and non-dominant hands with difficulty levels
nlongi_data.HV.doml1 <- subset(nlongi_data.HV, (dominant_hand_use=="YES") & difficulty_level==1)
nlongi_data.HV.doml2 <- subset(nlongi_data.HV, (dominant_hand_use=="YES") & difficulty_level==2)
nlongi_data.HV.doml3 <- subset(nlongi_data.HV, (dominant_hand_use=="YES") & difficulty_level==3)
nlongi_data.HV.ndoml1 <- subset(nlongi_data.HV, (dominant_hand_use=="NO") & difficulty_level==1)
nlongi_data.HV.ndoml2 <- subset(nlongi_data.HV, (dominant_hand_use=="NO") & difficulty_level==2)
nlongi_data.HV.ndoml3 <- subset(nlongi_data.HV, (dominant_hand_use=="NO") & difficulty_level==3)


#### Calculate ICC values
n.icc.HV.doml1 <- ICC_func(nlongi_data.HV.doml1)
n.icc.HV.doml2 <- ICC_func(nlongi_data.HV.doml2)
n.icc.HV.doml3 <- ICC_func(nlongi_data.HV.doml3)
n.icc.HV.ndoml1 <- ICC_func(nlongi_data.HV.ndoml1)
n.icc.HV.ndoml2 <- ICC_func(nlongi_data.HV.ndoml2)
n.icc.HV.ndoml3 <- ICC_func(nlongi_data.HV.ndoml3)


######################## Training set
## Difficulty Level 1
nfull_feat_train_doml1 <- newfull_feature_train %>% 
  subset(difficulty_level=="1" & dominant_hand_use=="YES") 
nfull_feat_train_ndoml1 <- newfull_feature_train %>% 
  subset(difficulty_level=="1" & dominant_hand_use=="NO") 

## Difficulty Level 2
nfull_feat_train_doml2 <- newfull_feature_train %>% 
  subset(difficulty_level=="2" & dominant_hand_use=="YES") 
nfull_feat_train_ndoml2 <- newfull_feature_train %>% 
  subset(difficulty_level=="2" & dominant_hand_use=="NO")

## Difficulty Level 3
nfull_feat_train_doml3 <- newfull_feature_train %>% 
  subset(difficulty_level=="3" & dominant_hand_use=="YES") 
nfull_feat_train_ndoml3 <- newfull_feature_train %>% 
  subset(difficulty_level=="3" & dominant_hand_use=="NO") 

######################## Test set
## Difficulty Level 1
nfull_feat_test_doml1 <- newfull_feature_test %>% 
  subset(difficulty_level=="1" & dominant_hand_use=="YES") 
nfull_feat_test_ndoml1 <- newfull_feature_test %>% 
  subset(difficulty_level=="1" & dominant_hand_use=="NO") 

## Difficulty Level 2
nfull_feat_test_doml2 <- newfull_feature_test %>% 
  subset(difficulty_level=="2" & dominant_hand_use=="YES") 
nfull_feat_test_ndoml2 <- newfull_feature_test %>% 
  subset(difficulty_level=="2" & dominant_hand_use=="NO")

## Difficulty Level 3
nfull_feat_test_doml3 <- newfull_feature_test %>% 
  subset(difficulty_level=="3" & dominant_hand_use=="YES") 
nfull_feat_test_ndoml3 <- newfull_feature_test %>% 
  subset(difficulty_level=="3" & dominant_hand_use=="NO") 

```



```{r  message=FALSE, warning=FALSE, include=FALSE}

nF <- length(grep("F",names(newfull_feature_train),value = TRUE)) ## number of features

########################################## Training sets
##### Statistical Significant Features
## DL1
nfull_feat_train_doml1.Sign <- nfull_feat_train_doml1 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_doml1, 0.001, "diagnosis_group", "F")))
nfull_feat_train_ndoml1.Sign <- nfull_feat_train_ndoml1 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_ndoml1, 0.001, "diagnosis_group", "F")))

## DL2
nfull_feat_train_doml2.Sign <- nfull_feat_train_doml2 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_doml2, 0.001, "diagnosis_group", "F")))
nfull_feat_train_ndoml2.Sign <- nfull_feat_train_ndoml2 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_ndoml2, 0.001, "diagnosis_group", "F")))

## DL3
nfull_feat_train_doml3.Sign <- nfull_feat_train_doml3 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_doml3, 0.001, "diagnosis_group", "F")))
nfull_feat_train_ndoml3.Sign <- nfull_feat_train_ndoml3 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_train_ndoml3, 0.001, "diagnosis_group", "F")))

###############
## Compute Fold-Change by HV vs. MS
nfc.doml1 <- HV.v.MS_func(nfull_feat_train_doml1.Sign, "YES", "1","diagnosis_group","F")
nfc.doml2 <- HV.v.MS_func(nfull_feat_train_doml2.Sign, "YES", "2","diagnosis_group","F")
nfc.doml3 <- HV.v.MS_func(nfull_feat_train_doml3.Sign, "YES", "3","diagnosis_group","F")

nfc.ndoml1 <- HV.v.MS_func(nfull_feat_train_ndoml1.Sign, "NO", "1","diagnosis_group","F")
nfc.ndoml2 <- HV.v.MS_func(nfull_feat_train_ndoml2.Sign, "NO", "2","diagnosis_group","F")
nfc.ndoml3 <- HV.v.MS_func(nfull_feat_train_ndoml3.Sign, "NO", "3","diagnosis_group","F")

#### Combine Fold-change and ICC of healthy volunteers data frame
n.icc.fc.doml1 <- data.frame(merge(n.icc.HV.doml1, nfc.doml1, by = "Feature"), 
                             difficulty = "Difficulty Level 1")
n.icc.fc.doml2 <- data.frame(merge(n.icc.HV.doml2, nfc.doml2, by = "Feature"), 
                             difficulty = "Difficulty Level 2")
n.icc.fc.doml3 <- data.frame(merge(n.icc.HV.doml3, nfc.doml3, by = "Feature"), 
                             difficulty = "Difficulty Level 3")

n.icc.fc.ndoml1 <- data.frame(merge(n.icc.HV.ndoml1, nfc.ndoml1, by = "Feature"), 
                              difficulty = "Difficulty Level 1")
n.icc.fc.ndoml2 <- data.frame(merge(n.icc.HV.ndoml2, nfc.ndoml2, by = "Feature"), 
                              difficulty = "Difficulty Level 2")
n.icc.fc.ndoml3 <- data.frame(merge(n.icc.HV.ndoml3, nfc.ndoml3, by = "Feature"), 
                              difficulty = "Difficulty Level 3")

n.icc.fc.dom <- data.frame(rbind(n.icc.fc.doml1, n.icc.fc.doml2, n.icc.fc.doml3), hand = "Dominant")
n.icc.fc.ndom <- data.frame(rbind(n.icc.fc.ndoml1, n.icc.fc.ndoml2, n.icc.fc.ndoml3), hand = "Non-dominant")

n.icc.fc <- rbind(n.icc.fc.dom,n.icc.fc.ndom) ### Merge the entire datasets of icc vs. fold change

############################################### Test sets

##### Statistical Significant Features
## DL1
nfull_feat_test_doml1.Sign <- nfull_feat_test_doml1 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_doml1, 0.001, "diagnosis_group", "F")))
nfull_feat_test_ndoml1.Sign <- nfull_feat_test_ndoml1 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_ndoml1, 0.001, "diagnosis_group", "F")))

## DL2
nfull_feat_test_doml2.Sign <- nfull_feat_test_doml2 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_doml2, 0.001, "diagnosis_group", "F")))
nfull_feat_test_ndoml2.Sign <- nfull_feat_test_ndoml2 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_ndoml2, 0.001, "diagnosis_group", "F")))

## DL3
nfull_feat_test_doml3.Sign <- nfull_feat_test_doml3 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_doml3, 0.001, "diagnosis_group", "F")))
nfull_feat_test_ndoml3.Sign <- nfull_feat_test_ndoml3 %>% 
  dplyr::select(-setdiff(paste("F",1:nF,sep = ""), wtest_func_onedl(nfull_feat_test_ndoml3, 0.001, "diagnosis_group", "F")))

###############

## Compute Fold-Change by HV vs. MS
nfc.doml1.tst <- HV.v.MS_func(nfull_feat_test_doml1.Sign, "YES", "1","diagnosis_group","F")
nfc.doml2.tst <- HV.v.MS_func(nfull_feat_test_doml2.Sign, "YES", "2","diagnosis_group","F")
nfc.doml3.tst <- HV.v.MS_func(nfull_feat_test_doml3.Sign, "YES", "3","diagnosis_group","F")

nfc.ndoml1.tst <- HV.v.MS_func(nfull_feat_test_ndoml1.Sign, "NO", "1","diagnosis_group","F")
nfc.ndoml2.tst <- HV.v.MS_func(nfull_feat_test_ndoml2.Sign, "NO", "2","diagnosis_group","F")
nfc.ndoml3.tst <- HV.v.MS_func(nfull_feat_test_ndoml3.Sign, "NO", "3","diagnosis_group","F")

#### Combine Fold-change and ICC of healthy volunteers data frame
n.icc.fc.doml1.tst <- data.frame(merge(n.icc.HV.doml1, nfc.doml1.tst, by = "Feature") ,
                               difficulty="Difficulty Level 1") 
n.icc.fc.doml2.tst <- data.frame(merge(n.icc.HV.doml2, nfc.doml2.tst, by = "Feature"),
                               difficulty="Difficulty Level 2")
n.icc.fc.doml3.tst <- data.frame(merge(n.icc.HV.doml3, nfc.doml3.tst, by = "Feature"),
                               difficulty="Difficulty Level 3")

n.icc.fc.ndoml1.tst <- data.frame(merge(n.icc.HV.ndoml1, nfc.ndoml1.tst, by = "Feature"),
                               difficulty="Difficulty Level 1")
n.icc.fc.ndoml2.tst <- data.frame(merge(n.icc.HV.ndoml2, nfc.ndoml2.tst, by = "Feature"),
                               difficulty="Difficulty Level 2")
n.icc.fc.ndoml3.tst <- data.frame(merge(n.icc.HV.ndoml3, nfc.ndoml3.tst, by = "Feature"),
                               difficulty="Difficulty Level 3")


n.icc.fc.dom.tst <- data.frame(rbind(n.icc.fc.doml1.tst, n.icc.fc.doml2.tst, n.icc.fc.doml3.tst), 
                             hand = "Dominant")
n.icc.fc.ndom.tst <- data.frame(rbind(n.icc.fc.ndoml1.tst, n.icc.fc.ndoml2.tst, 
                                    n.icc.fc.ndoml3.tst), hand = "Non-dominant")

n.icc.fc.tst <- rbind(n.icc.fc.dom,n.icc.fc.ndom) ### Merge the entire datasets of icc vs. fold change

```



No we plot the ICC vs the Log2 Fold Change for significant features

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}
##################### PLOTTING ICC vs. Log2FC

#### PLot in facet_grid
ggplot(n.icc.fc, aes(x=Log2FC, y=ICC, label = Label)) +
    geom_point(alpha=0.7, size = 8, color = "darkgray") +
    geom_text(size = 5) +
    labs(x = "", y ="") +
    facet_grid(hand~difficulty, scales = "free") +
    theme_bw() + Nice.Label


ggplot(n.icc.fc, aes(x=Log2FC, y=ICC, label = Label)) +
    geom_point(alpha=0.7, size = 8, color = "darkgray") +
    geom_text(size = 5) + xlim(1,4.5)+ylim(0.5,0.75)+
    labs(x = "", y ="") +
    facet_grid(hand~difficulty) +
    theme_bw() + Nice.Label


```


## Correlation plots

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}

#### Correlation plots
## Training set
## Difficulty Level 1
mat.feat_func.onel2(nfull_feat_train_doml1,new_cl.HV.MS.S,"MS",0.05)
mat.feat_func.onel2(nfull_feat_train_ndoml1,new_cl.HV.MS.S,"MS",0.05)
mat.feat_func.onel2(nfull_feat_test_doml1,new_cl.HV.MS.S,"MS",0.05)
mat.feat_func.onel2(nfull_feat_test_ndoml1,new_cl.HV.MS.S,"MS",0.05)

```


## Correlation between Hausdorff Error and time to complete drawings

Now we will plot Sum of hausdorff Distance (F24) vs. time taken to complete drawings (F35) 


```{r  message=FALSE, warning=FALSE, include=FALSE}
## Create the data to plot the F24 (sum of Hausdorff distance) vs. time taken to complete drawing (F35)

##### Training set
## Healthy Cohorts
F24F35_train.HV <- full_feature_train %>% 
  dplyr::select(difficulty_level, diagnosis_group, dominant_hand_use, F24, F35) %>% 
  dplyr::filter(diagnosis_group=="HV") %>% dplyr::select(-diagnosis_group) %>% 
  dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
  dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 

## MS Cohorts
F24F35_train.MS <- full_feature_train %>% 
  dplyr::select(difficulty_level, diagnosis_group, dominant_hand_use, F24, F35) %>% 
  dplyr::filter(diagnosis_group=="MS") %>% dplyr::select(-diagnosis_group) %>% 
  dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
  dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 
  

##### Test set
## Healthy Cohorts
F24F35_test.HV <- full_feature_test2 %>% 
  dplyr::select(difficulty_level, diagnosis_group, dominant_hand_use, F24, F35) %>% 
  dplyr::filter(diagnosis_group=="HV") %>% dplyr::select(-diagnosis_group) %>% 
  dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
  dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 

## MS Cohorts
F24F35_test.MS <- full_feature_test2 %>% 
  dplyr::select(difficulty_level, diagnosis_group, dominant_hand_use, F24, F35) %>% 
  dplyr::filter(diagnosis_group=="MS") %>% dplyr::select(-diagnosis_group) %>% 
  dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
  dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 

```




```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}

### Create a function to extract correlation coefficients
### Equation for the correlation coefficient
corfunc <- function(df){
        #setting the correlation function. 
        cor_fun <-cor.test(df$F35,df$F24, method = "spearman") #regression function
        #getting the pvalue and correlation coefficient
        Rho <-round(cor_fun$estimate,3)  
        pval <- ifelse(cor_fun$p.value < 0.001,"< 0.001",round(cor_fun$p.value,3)) 
        c(Rho,pval)
}

### Extract the coefficient of the regression lines

cor_F24F35_train.HV <- plyr::ddply(F24F35_train.HV,c("difficulty","hand"),corfunc)
colnames(cor_F24F35_train.HV) <- c("difficulty","hand","rho","pvalue")

cor_F24F35_train.MS <- plyr::ddply(F24F35_train.MS,c("difficulty","hand"),corfunc)
colnames(cor_F24F35_train.MS) <- c("difficulty","hand","rho","pvalue")

cor_F24F35_test.MS <- plyr::ddply(F24F35_test.MS,c("difficulty","hand"),corfunc)
colnames(cor_F24F35_test.MS) <- c("difficulty","hand","rho","pvalue")


##################### PLOTTING F24 vs. F35

#### PLot in facet_grid
##### Training set
## HC
ggplot(F24F35_train.HV, aes(x=F24/1000, y=F35)) +
     geom_point(alpha=1, size = 2, color = "black") +
     geom_smooth(method=lm, color="blue", fill="darkgray") + ## regression line with 95% confidence interval
     labs(x = "", y ="") +
     facet_grid(hand~difficulty, scales = "free") +
     geom_label(data=cor_F24F35_train.HV, inherit.aes=FALSE, 
                aes(x = c(2,2,5,5,25,25), y = c(25,30,25,30,25,30),
                    label= paste("ρ =",rho,","," ","P",pvalue)
                ), label.size = 1, size = 5) +
     theme_bw() + Nice.Label

## MS
ggplot(F24F35_train.MS, aes(x=F24/1000, y=F35)) +
    geom_point(alpha=1, size = 2, color = "black") +
    geom_smooth(method=lm, color="blue", fill="darkgray") + ## regression line with 95% confidence interval
    labs(x = "", y ="") +
    facet_grid(hand~difficulty, scales = "free") +
    geom_label(data=cor_F24F35_train.MS, inherit.aes=FALSE, 
             aes(x = c(20,20,45,45,90,90), y = c(75,75,75,75,75,75),
                 label= paste("ρ =",rho,","," ","P",pvalue)
                 ), label.size = 1, size = 5) +
    theme_bw() + Nice.Label

##### Test set
## MS

ggplot(F24F35_test.MS, aes(x=F24/1000, y=F35)) +
    geom_point(alpha=1, size = 2, color = "black") +
    geom_smooth(method=lm, color="blue", fill="darkgray") + ## regression line with 95% confidence interval
    labs(x = "", y ="") +
    facet_grid(hand~difficulty, scales = "free") +
    geom_label(data=cor_F24F35_test.MS, inherit.aes=FALSE, 
             aes(x = c(10,10,20,20,50,50), y = c(35,35,35,35,35,35),
                 label= paste("ρ =",rho,","," ","P",pvalue)
                 ), label.size = 1, size = 5) +
    theme_bw() + Nice.Label

```


## Linear regression between Hausdorff Error and clinical features



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}


### Create a function to extract regression coefficient
### Equation for the linear regression model
regression <- function(df){
  # split the data ionto training and test set
  train <- df %>% subset(data_type=="train") %>% dplyr::select(-c(data_type,cl))
  test <- df %>% subset(data_type=="test") %>% dplyr::select(-c(data_type,cl))
  
  cl <- as.character(df$cl[1]) ## extract the clinical feature name
  #setting the regression function. 
  reg_fit <-lm(formula = train[,cl]~F24, data = train) # linear regression fit 
  
  # Getting the R square and pvalue of the model fit 
  R2 <- round(summary(reg_fit)$r.squared, 4)
  pvalue <- ifelse(anova(reg_fit)$'Pr(>F)'[1] < 0.001, "< 0.001",round(anova(reg_fit)$'Pr(>F)'[1],4))
  
  # Getting the RMSE and R square using the test set
  ## Make prediction on the test data
  predictions <- reg_fit %>% predict(data.frame(F24=test$F24))
  
  ## Compute the average prediction error RMSE
  rmse_test <- RMSE(predictions, test[,cl])
  r2_test <- R2(predictions, test[,cl])
  
  c(R2, pvalue, rmse_test, r2_test)
}


#### Plot regression between the four clinical features and F24

lm.regression_func <- function(fdata_train, fdata_test, cdata, diagnosis, cl_feat){
  ## fdata is the daframe containing the F1, F9, F13, and F24 features
  ## cdata is the clinical features data
  ## cl_feat is the clinical feature we will use as ground truth in quotation
  ## diagnosis is diagnosis group (HV or MS)
  
  ## Take log of f24 and C1
  train_data <- fdata_train %>% dplyr::mutate(F24 = bisym.log_func(F24))
  test_data <- fdata_test %>% dplyr::mutate(F24 = bisym.log_func(F24))
  cl_data <- cdata %>% dplyr::mutate(C1 = bisym.log_func(C1)) %>% 
    dplyr::mutate(C4 = bisym.log_func(C4)) %>% 
    dplyr::mutate(C5 = bisym.log_func(C5)) %>% 
    dplyr::mutate(C6 = bisym.log_func(C6))
  
  ## Remove outliers and select diagnosis group
  train_data2 <- train_data %>% #ungroup() %>% group_by(difficulty_level, dominant_hand_use) %>%
    group_modify(~outlier.df_func(.x)) %>% dplyr::filter(diagnosis_group==diagnosis)
  test_data2 <- test_data %>% #ungroup() %>% group_by(difficulty_level, dominant_hand_use) %>%
    group_modify(~outlier.df_func(.x)) %>% dplyr::filter(diagnosis_group==diagnosis)
  
  
  ## return all rows from pat_data where there are matching in cl_data 
  merge.train_data <- dplyr::inner_join(train_data2, cl_data, by =
                                          c("patient_id","testdate","diagnosis_group"))
  merge.test_data <- dplyr::inner_join(test_data2, cl_data, by =
                                         c("patient_id","testdate","diagnosis_group"))
  
  ## Extract the training F24 and y=Clinical datasets
  train_sub <- merge.train_data %>% ungroup() %>% 
    dplyr::select(difficulty_level, dominant_hand_use, cl_feat, F24) %>% 
    dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
    dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                  hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                  difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                  difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                  difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 
  
  ## Extract the test X and y datasets
  test_sub <- merge.test_data %>% ungroup() %>% 
    dplyr::select(difficulty_level, dominant_hand_use, cl_feat, F24) %>% 
    dplyr::rename(hand=dominant_hand_use, difficulty=difficulty_level) %>% 
    dplyr::mutate(hand=replace(hand,hand=="YES","Dominant"), 
                  hand=replace(hand,hand=="NO","Non-dominant"), difficulty = as.character(difficulty),
                  difficulty=replace(difficulty,difficulty=="1","Difficulty Level 1"),
                  difficulty=replace(difficulty,difficulty=="2","Difficulty Level 2"),
                  difficulty=replace(difficulty,difficulty=="3","Difficulty Level 3")) 
  
  ###### Extract the coefficient of the regression lines
  ## combine the training and test set
  all_data <- rbind(data.frame(train_sub, data_type="train",cl=cl_feat),
                    data.frame(test_sub, data_type="test",cl=cl_feat))
  
  reg_coeff <- plyr::ddply(all_data, c("difficulty","hand"), regression)
  colnames(reg_coeff) <- c("difficulty","hand","R2_train","pval_train","RMSE_test","R2_test")
  # 
  # 
  p <- ggplot(train_sub, aes(x=F24, y= C1)) + ## manually change the clinical feature before running
    geom_point(alpha=1, size = 2, color = "black") +
    geom_smooth(method=lm, color="blue", fill="darkgray") + ## regression line with 95% confidence interval
    labs(x = "", y ="") +
    facet_grid(hand~difficulty, scales = "free") +
    geom_label(data=reg_coeff, inherit.aes=FALSE,
               aes(x = c(3.5,3.5,4,4,4.1,4.1), y = c(1.25,1.25,1.25,1.25,1.25,1.25),## change adequately
                   label= paste("R-sq. = ",R2_train,","," ","P = ",pval_train)
               ), label.size = 1, size = 5) + #ylim(0.0,2.8) +
    theme_bw() + Nice.Label
  
  return(list(p,reg_coeff))
  
}




```



```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=12, fig.width= 16}
## C1, C4, C5, and C6 are the clinical features of interest
lm.regression_func(full_feature_train, full_feature_test2, new_cl.HV.MS, "HV", "C1")

lm.regression_func(full_feature_train, full_feature_test2, new_cl.HV.MS, "MS", "C1")

```





